### 2016-12-28 Paper2/BOAST.tex
section boast

\documentclass{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}

\newcommand{\productname}[1]{\textsc{#1}}
\newcommand{\latin}[1]{\textit{#1}}
\newcommand{\eg}{\latin{e.g.,}\xspace}
\newcommand{\etc}[1]{\latin{etc...}}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand*{\Num}{N\textsuperscript{o}\xspace}

%\usepackage[textwidth=17mm]{todonotes}
\usepackage{todonotes}

\newcommand{\FD}[2]{\todo[inline, size=\small]{\textbf{FD$\rightarrow$#1}: #2.}}

\lstdefinestyle{customc}{
  language=C,
  basicstyle=\scriptsize,
  backgroundcolor=\color{white},
  frame=single,
  captionpos=b,
  morekeywords={int64_t, int32_t}
}

\lstdefinestyle{customgputrace}{
  language=C,
  basicstyle=\scriptsize,
  backgroundcolor=\color{white},
  frame=single,
  captionpos=b,
  morekeywords={New, kernel, buffer, Mb, READ_WRITE, Buffer, written,
    b, at, out}
}

\lstdefinestyle{customcl}{
  language=C,
  basicstyle=\scriptsize,
  backgroundcolor=\color{white},
  frame=single,
  captionpos=b,
  morekeywords={__global, __local, global, local, __kernel, kernel, short,
uchar, uchar16, int16, short16}
}

\lstdefinestyle{BOAST}{
  language=Ruby, 
  basicstyle=\scriptsize, 
  backgroundcolor=\color{white}, 
  frame=single, 
  captionpos=b,
  morekeywords={BOAST, pr, decl, opn, close}
}

\lstdefinestyle{customf}{
  language=FORTRAN,
  basicstyle=\scriptsize,
  backgroundcolor=\color{white},
  frame=single,
  captionpos=b
}

\lstdefinestyle{custombash}{
  language=Bash,
  basicstyle=\scriptsize,
  backgroundcolor=\color{white},
  frame=single,
  captionpos=b
}

\usepackage{url}

\bibliographystyle{IEEEtran}

\begin{document}

\title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}

%% or include affiliations in footnotes:
\author{
   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{2}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{2}Centre National de la Recherche Scientifique}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Joseph Fourier}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/SP2M/L\_Sim}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}
}

\markboth{International Journal of Scientific Computing}
{Videau \MakeLowercase{\textit{et al.}}: BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
%\address[mymainaddress]{LIG/CNRS}

%\begin{keyword}
%  Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non Regression Testing \sep Productivity and Software Design \sep High Performance Computing
%\end{keyword}

%\end{frontmatter}
\maketitle

\begin{abstract}
The portability of real HPC applications on new platforms is an open and very
delicate problem.  Especially the performance portability of the underlying
computing kernels is problematic as they need to be tuned for each and every
platform the application encounters. This paper presents BOAST, a
metaprogramming framework dedicated to computing kernels.  BOAST allows the
description of a kernel and its possible optimizations using a DSL. BOAST
runtime will then compare the different versions' performance as well as verify
their exactness. BOAST is applied to three use cases: a Laplace kernel in
OpenCL and two HPC applications BigDFT (electronic density computation) and
SPECFEM3D (seismic and wave propagation).
\end{abstract}

%%% INTRODUCTION %%%
\section{Introduction}

Porting and tuning HPC applications to new platforms is tedious and costly in
term of human resources.  Many developer teams are more concentrated on solving
the physical problems faster and better with new algorithms and simulating new
physics. This is why optimization, while important, is a task which is
sometimes neglected or hidden for many developers.  In the same way, Domain
Specific Language (DSLs) have exactly this goal, the complete separation of the
high-level code dedicated to simulate new physics and the low-level one which
is more dependent on the HPC architecture. Nevertheless, decoupling the high
and low levels does not avoid the need of optimizations which become more and
more unavoidable with new architectures based on many cores and accelerators.

Unfortunately, portability efforts of the low level codes (i.e. computing
kernels) are often lost when migrating to a new architecture. Worse, code may
lose maintainability because several versions of some functionalities coexist,
usually with a lot of duplication.  Thus productivity of porting and tuning
efforts is low as a huge fraction of those developments are never used after
the platform they were intended for is decommissioned.

The primary goal of the BOAST metaprogramming framework is to simplify the
optimization of computing kernels for HPC applications. The main idea is to use
a high level language to define an API of computing kernels i.e. subroutines
and their related arguments, but also describe the way to optimize it
(unrolling, vectorization, parametrization, templating, tiling). BOAST will
then test automatically the many possibilities of optimizations and choose the
best one after validation from a reference version. Then BOAST can generate
pieces of Fortran, C, OpenCL, or CUDA codes, test them for a specific target
architecture.

With the definition of an API, BOAST can hide the implementation details at
high level but also is a way to factorize different kernels depending only on
parameters such as data types, length of filters, or boundary conditions. The
maintainability and genericity of HPC codes is often limited. One of the reason
is that producing generic code in Fortran 90/95 is difficult as the language
does not really fit for it. Sometimes, adding genericity or functionalities can
be quite costly and may degrade performance as optimization opportunities that
come from over-specification are lost. BOAST can provide a solution to handle
genericity and add new functionalities at low levels.

The paper is organized as follows. First in Section~\ref{sec:backg} we further
motivate the need for metaprogramming frameworks.  Secondly in
Section~\ref{sec:laplace} we illustrate the optimization problems related to
scientific computing with the Laplace equation. Thirdly we present our
solution, called BOAST, its architecture and its functionalities in
Section~\ref{sec:boast}. Fourthly we apply BOAST to three use cases: the
Laplace equation motivating example, short multi-dimensional convolutions from
the BigDFT quantum chemistry application, and the seismic wave propagation
simulator SPECFEM3D where 80 GPU computing kernels have been ported improving
considerably the maintainability and providing a single code base for OpenCL
and CUDA kernel versions. Finally, before a conclusion and future works, we
compare BOAST with other related works in Section~\ref{sec:related}.

%%% BACKGROUND AND MOTIVATION %%%
\section{Background and Motivation\label{sec:backg}}

The motivation of the BOAST project comes from the need of the HPC communities
which have full and large applications.  These applications are difficult to
port and maintain for emerging architectures especially since new architectures
are based on accelerators, longer vector units, or low power processors.

%%% EVOLUTION OF HPC ARCHITECTURES %%%
\subsection{Evolution of HPC Architectures}

Evolution of HPC Architectures is rapid and also diverse. In the last seven
years no less than 7 architectures have been number one in the
Top500~\cite{webtop500}:
\begin{itemize}
\item Sunway processor (TaihuLight, 2016)
\item Intel Processor + Xeon Phi (Tianhe-2, 2013)
\item AMD Processor + NVIDIA GPU (Titan, 2012)
\item IBM BlueGene/Q (Sequoia, 2012)
\item Fujitsu SPARC64 (K computer, 2011)
\item Intel Processor + NVIDIA GPU (Tianhe-1, 2010)
\item AMD Processor (Jaguar, 2009)
\end{itemize}

Being able to efficiently use those architectures on such a small period is
challenging for scientists and application developers.

The race to exascale is not going to simplify the environment. All of the above
machine architectures can be considered to build these supercomputers and other
will appear soon. For instance European FP7 project DEEP considers using
Accelerators (XEON Phi) while the European FP7 project Mont-Blanc considers
using low-power embedded processor with integrated GPU.

Running existing applications on these new architectures is an open research
subject as well as an ongoing porting effort.  Thus, those projects have work
packages dedicated to applications. Those work packages are dedicated to
porting and optimizing Scientific applications on those new architectures. In
the DEEP project six applications were selected for porting and optimizing,
eleven were selected in the Mont-Blanc project.

%%% SCIENTIFIC COMPUTING APPLICATIONS %%%
\subsection{Scientific Computing Applications}

Scientific Computing Applications are usually developed by other science
researchers such as physicists, chemists or meteorologists. Those codes are
usually written in Fortran for historical and performance reasons. Codes can be
quite huge (several thousands Lines of Code, LoC) with lots of functionalities.
Nonetheless, they are usually based on computing kernels.  Computing kernels
are resource intensive and well-defined parts of a program that usually work on
precisely defined data. Those kernels represent the most time-consuming part of
an HPC application, and consequently, they are the prime target for
optimizations.

Those applications are often developed by several individuals. Sometimes, some
of those developers only work a few months on the application. Maintaining
optimized code written by someone else is quite a challenge. Several languages
and programming paradigms can also be used in a project and thus the maintainer
must be knowledgeable in several areas of expertise. Portability problems can
also be caused by the availability of an optimizing compiler for a specific
language and architecture. C compilers are usually first available while
Fortran may sometimes come later.

In Section~\ref{use_cases} we will present two HPC applications that we used as
use cases: SPECFEM3D and BigDFT. They are both based on computing kernels and
were selected as candidate applications in the Mont-Blanc project.

%%% HOW SHOULD COMPUTING KERNELS BE WRITTEN? %%%
\subsection{How Should Computing Kernels be Written?}

The problematic here is to obtain computing kernels that present good
performance on the architectures encountered by the application while staying
portable after the optimization process took place. Indeed the application
might encounter one of the many architectures that can be found in HPC.
Investing manpower to optimize the application for a new architecture is
reasonable, suffering hindrance from previous optimization work is not. Thus
optimizations have to be as orthogonal as possible from one another so as to be
easily activated and deactivated.

If this paradigm is followed by developers then they will rapidly be confronted
with a huge optimization space to search. They will need to be able to test
easily the performance impact of the chosen optimizations without running the
full application. The same reasoning implies that kernels should be tested for
non regression without running the full application.

What we want is computing kernels that are written in a \emph{portable} manner,
in a way that raises developer \emph{productivity}, and they must present good
\emph{performance}.

%%% BASIC EXAMPLE: OPENCL LAPLACE %%%
\section{Basic Example: OpenCL Laplace\label{sec:laplace}}

During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
on the Mali GPU was given. One of the case study was a Laplace
filter~\cite{opencl_arm_training}. This is a good example of an algorithm that
is simple in its formulation but can be complex to optimize because many
different optimizations are available and can be combined together. The correct
combination of optimizations will depend on the target architecture.

%%% THE LAPLACE FILTER %%%
\subsection{The Laplace Filter}

Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace
filter in C99. For sake of clarity, boundary conditions have been omitted.

\lstset{style=customc,
        caption={Laplace C99 Filter.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_c_filter}}
\begin{lstlisting}
void laplace(const int width, const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp =
   -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
  - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
  - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp<0 ? 0 : (tmp>255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}

A naive implementation in OpenCL is proposed in
Listing~\ref{lst:laplace_opencl_filter}. Each work item processes one pixel of
the resulting image.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter}}
\begin{lstlisting}
kernel laplace(const int width, const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);
  for (int c = 0; c < 3; c++) {
    int tmp = -src[3*width*(j-1) + 3*(i-1) + c]\
             - src[3*width*(j-1) + 3*(i  ) + c]\
             - src[3*width*(j-1) + 3*(i+1) + c]\
             - src[3*width*(j  ) + 3*(i-1) + c]\
           + 9*src[3*width*(j  ) + 3*(i  ) + c]\
             - src[3*width*(j  ) + 3*(i+1) + c]\
             - src[3*width*(j+1) + 3*(i-1) + c]\
             - src[3*width*(j+1) + 3*(i  ) + c]\
             - src[3*width*(j+1) + 3*(i+1) + c];
    dst[3*width*j + 3*i + c] = clamp(tmp, 0, 255);
  }
}
\end{lstlisting}

%%% POSSIBLE OPTIMIZATIONS OF THE OPENCL VERSION %%%
\subsection{Possible Optimizations of the OpenCL Version}

Several optimizations were proposed and successively applied to the previous
implementation.

\paragraph{Vectorization} The first proposed optimization involves computing
five pixels instead of one using the vectors available on the Mali
architecture. The vectors have to be kept in memory to ensure an efficient
execution. Using vectors of 16 elements, 15 useful components are
simultaneously loaded and can be used in computation.  This optimization yields
speedups between 1.5 and 6 depending on the image size.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector}}
\begin{lstlisting}
kernel laplace(const int width, const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);

  uchar16 v11_ =vload16(0,src + 3*width*(j-1) + 3*5*i -3);
  uchar16 v12_ =vload16(0,src + 3*width*(j-1) + 3*5*i   );
  uchar16 v13_ =vload16(0,src + 3*width*(j-1) + 3*5*i +3);
  uchar16 v21_ =vload16(0,src + 3*width*(j  ) + 3*5*i -3);
  uchar16 v22_ =vload16(0,src + 3*width*(j  ) + 3*5*i   );
  uchar16 v23_ =vload16(0,src + 3*width*(j  ) + 3*5*i +3);
  uchar16 v31_ =vload16(0,src + 3*width*(j+1) + 3*5*i -3);
  uchar16 v32_ =vload16(0,src + 3*width*(j+1) + 3*5*i   );
  uchar16 v33_ =vload16(0,src + 3*width*(j+1) + 3*5*i +3);

  int16 v11 = convert_int16(v11_);
  int16 v12 = convert_int16(v12_);
  int16 v13 = convert_int16(v13_);
  int16 v21 = convert_int16(v21_);
  int16 v22 = convert_int16(v22_);
  int16 v23 = convert_int16(v23_);
  int16 v31 = convert_int16(v31_);
  int16 v32 = convert_int16(v32_);
  int16 v33 = convert_int16(v33_);

  int16 res = v22 * (int)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
        res = clamp(res, (int16)0, (int16)255);
  uchar res_ = convert_uchar16(res);

  vstore8(res_.s01234567, 0, dst + 3*width*j + 3*5*i);
  vstore8(res_.s89ab,     0, dst + 3*width*j + 3*5*i + 8);
  vstore8(res_.scd,       0, dst + 3*width*j + 3*5*i + 12);
  dst[3*width*j + 3*5*i + 14] = res_.se;
}
\end{lstlisting}

\paragraph{Synthesizing Loads} It is possible to reduce the number of loads
since the vectors are overlapping.
Listing~\ref{lst:laplace_opencl_filter_vector_synth} shows how the vectors are
loaded in this case. This optimization yields marginal improvements.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector_synth}}
\begin{lstlisting}
  uchar16 v11_ =vload16(0, src + 3*width*(j-1) + 3*5*i -3);
  uchar16 v13_ =vload16(0, src + 3*width*(j-1) + 3*5*i +3);
  uchar16 v12_ =uchar16(v11_.s3456789a, v13_.s56789abc);
  uchar16 v21_ =vload16(0, src + 3*width*(j  ) + 3*5*i -3);
  uchar16 v23_ =vload16(0, src + 3*width*(j  ) + 3*5*i +3);
  uchar16 v22_ =uchar16(v21_.s3456789a, v23_.s56789abc);
  uchar16 v31_ =vload16(0, src + 3*width*(j+1) + 3*5*i -3);
  uchar16 v33_ =vload16(0, src + 3*width*(j+1) + 3*5*i +3);
  uchar16 v32_ =uchar16(v31_.s3456789a, v33_.s56789abc);
\end{lstlisting}

\paragraph{Temporary Variables Size} Using \emph{int} to store intermediary
results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
the code is modified to use smaller types. This yields a speedup of 1.3 for
most image sizes.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized Using \emph{short}.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_short}}
\begin{lstlisting}
  short16 v11 = convert_short16(v11_);
  short16 v12 = convert_short16(v12_);
  short16 v13 = convert_short16(v13_);
  short16 v21 = convert_short16(v21_);
  short16 v22 = convert_short16(v22_);
  short16 v23 = convert_short16(v23_);
  short16 v31 = convert_short16(v31_);
  short16 v32 = convert_short16(v32_);
  short16 v33 = convert_short16(v33_);

  short16 res = v22 * (short)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
          res = clamp(res, (short16)0, (short16)255);
\end{lstlisting}

\paragraph{More Optimizations} Several other optimizations can be attempted at
this point. For instance reducing or increasing the number of pixels each work
item is processing. They can yield improved performance in some cases,
especially when avoiding memory alignment problems. These optimizations will
not be shown here but can be found in~\cite{opencl_arm_training}. Others
consist in tiling the problem space or improving the loads synthesis.

%%% Improving the Methodology %%%
\subsection{Improving the Methodology}

The process described in the previous Subsection is quite tedious and requires
some intimate knowledge of the target architecture. The results are impressive,
as speedups of almost 10 are observed compared to the naive OpenCL
implementation.

Nonetheless, the process is frustrating. The optimizations are never evaluated
independently from one another. Some were arbitrarily configured (like the
number of pixels chosen for instance). Testing the different combinations of
those optimizations and the different parameters would be very costly in
developer time and, moreover, like all repetitive and tedious tasks, error
prone. Thus every created kernel would have to be thoroughly tested to ensure
no error was done.

In the next Section we will present our answer to these issues: BOAST.  BOAST
is a framework dedicated to kernel description, optimization, regression
testing, and autotuning.

%%% BOAST: USING CODE GENERATION IN APPLICATION AUTOTUNING %%%
\section{BOAST: Using Code Generation in Application Autotuning\label{sec:boast}}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{BOAST_Workflow.pdf}
\caption{Structure and Workflow of the BOAST framework.}
\label{fig:boast_workflow}
\end{center}
\end{figure}

BOAST provides scientific application developers with a framework to develop and
test application computing kernels~\cite{videau2013boast}.

Figure~\ref{fig:boast_workflow} illustrates the envisioned workflow and program
structure. The user starts from an application kernel (either designed or
implemented), and writes it in a dedicated language (Step 1). The language
provides enough flexibility for the kernel to be meta-programmed with several
orthogonal optimizations. Then the developer selects a kernel configuration and
a target language. Those parameters define the output source code that will be
generated by BOAST (Step 2). The resulting code source is then built according
to the user specified compiler and options (Step 3). If input data are available
then the kernel can be benchmarked and tested for non regression. Based on the
results, other optimizations can be selected (Step 4) or new optimizations can
be added to the BOAST sources. The process can be repeated until a good
candidate is found on the target platform. The resulting kernel is then added to
the program (Step 5).

Several steps in the workflow can be automated: optimization and compiler flags
exploration, non regression testing, benchmarking, \ldots This automation can be
scripted by the user or by interfacing other dedicated tools. In order
to achieve those goals three aspects should be considered: code description,
code generation, and kernel execution run-time.

%%% KERNEL DESCRIPTION LANGUAGE %%%
\subsection{Kernel Description Language}

Usually computing kernels are hotspots of an HPC application, and they take most
of the time based on loop nests. A lot of efforts are dedicated to their tuning
and the obtained result is often quite different from the original procedure.
Several transformations can be applied to such kernels. And those optimizations
are often applied manually as compiler may fail to recognise the opportunity.

There are many different loop optimization techniques~\cite{wolf1991loop}. We
can cite loop skewing~\cite{wolfe1986loops} (derives nested loops wavefronts)
or loop interchange~\cite{allen1984automatic} (loop variables change places).

\FD{BV}{Bizarre la suite ici. Ca fait plus ``previous work''}
The importance of  correct loop imbrication on BLAS~\cite{lawson1979basic}
operations is studied in~\cite{soliman2009performance}, and shows performance
increase of a factor up to 5 when using correct loop imbrication. The
importance of code transformation is stressed in~\cite{ye2011porting}, where a
selection of GPU kernels are ported to CPU and optimized.

BOAST kernel description language should be able to express all these
optimizations. This gives us a set of constraints to implement in the language:
\begin{itemize}
\item Arbitrary number of variables have to be created and manipulated (types,
attributes, \ldots).
\item Procedures have to be abstracted (reunite Fortran and C like languages,
attributes, \ldots).
\item Functions must be available.
\item Variables, constants and functions should be able to be composed in
complex expressions.
\item Basic control structures (for, while, if/else, , \ldots) have to be abstracted.
\item Powerful array management features (allowing several dimensions, transformations,
indexing, \ldots).
\end{itemize}

In order to manipulate those abstractions we want to have a syntax similar to 
what programmers use. For instance, commonly used operators have to be
available and behave as expected. It must also be possible to differentiate an
action on an abstraction in the context of an expression and in the context of
the management of this expression. $c = a + b$, an expression that affects the
results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves
the expression $a + b$ to a variable $c$. This is why an embedded Domain
Specific Language approach was selected~\cite{hudak1996building}. This feature
allows for the coexistence of two languages: the host language and the Domain
Specific Language (DSL).  In our case, DSL allows the description of the kernel
($c = a + b$) while the host language provides the meta programming of the
kernel ($c \leftarrow a + b$). Operator overloading of the host language will
provide the familiar syntax programmers are accustomed to.

It was also important to have our constructs like \emph{for} loops to have a
syntax approaching those commonly found in programming languages. To this end,
we needed a language which could seamlessly pass a block of code to a function.
Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
capabilities as well. This is the main reason why it was selected for BOAST.

One of the added advantage of using a high-level scripting language like Ruby
as the host language if its interfacing capabilities, providing an easy way to
use the libraries needed during the development of the framework.

%%% BOAST KEYWORDS %%%
\subsubsection{BOAST Keywords}

In order to clearly differentiate what is going to be generated from what is
related to manipulations in the host language four keywords were defined. They
are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
Embedded Domain Specific Language (EDSL), these four keywords are methods in
the BOAST namespace. Sample usage of these keywords will be found in the next
Figures.
%Usage of these keywords can be found on Figure~\ref{fig:keywords}.

The \emph{decl} method is used to declare variables or procedures and functions. 
The \emph{pr} method calls the public \emph{pr} method of objects it is
called on. Each BOAST object is responsible for printing itself correctly
depending on the BOAST configuration at the time the print public method is
called. Calling directly the \emph{pr} method of a BOAST object yields the same
result. The \emph{opn} method can be used to print the beginning of a control
structure without an associated code block.
The \emph{close} method is the counterpart to the \emph{opn} method. It is used
to close a control structure without an associated code block.

%%% BOAST ABSTRACTIONS %%%
\subsubsection{BOAST Abstractions}

BOAST defines several classes that are used to represent the structure of the
code. These classes can be sorted in two groups, algebraic related and control
flow related.

%%% ALGEBRA %%%
\paragraph{Algebra}

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_algebra},
        numbers=left }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

\begin{minipage}[b]{0.5\linewidth}
\centering
\lstset{style=customf,
        caption={Fortran output.},
        label={lst:BOAST_algebra_FORTRAN} }

\begin{lstlisting}
integer(kind=4) :: i
integer(kind=8) :: k
real(kind=8), dimension(-5:21) :: l
i = 5
k = (i+5)*(2)
l(k) = 1.0_wp
l(k+1) = sin(i+5)
\end{lstlisting}
\end{minipage}
\hspace{0.08\linewidth}
\begin{minipage}[b]{0.40\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_algebra_C} }

\begin{lstlisting}
int32_t i;
int64_t k;
double l[27];
i = 5;
k = (i+5)*(2);
l[k-(-5)] = 1.0;
l[k+1-(-5)] = sin(i+5);
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for Variables and Expressions.}
\label{fig:BOAST_algebra}
\end{figure}
The first and most fundamental abstraction is named Variable. Variables have a
type, a name, and a set of named attributes. The existing attributes are mainly
inspired from Fortran. Those attributes are not limited and can be arbitrarily
enriched, allowing a lot of flexibility in Variable management.

The second abstraction is named Expression. It combines variables into
algebraic or logic expressions. Most of the classical operators are overloaded
for those two abstractions and thus the syntax of the expressions are rather
straightforward. The exception is the assignment operator as it is important to
differentiate between assigning an Expression or a Variable in the Ruby context
and the assignment operator in the context of a BOAST Expression. Thus the
assignment in a BOAST Expression is represented as the \emph{===} operator,
while the classical assignment is kept as the \emph{=} operator. Function calls
(FuncCall) are also abstracted and can be used in Expressions.

Figure~\ref{fig:BOAST_algebra} shows some basic usage of both Variables and
Expressions as well as the \emph{pr} and \emph{decl} keywords. For clarity we
stayed out of BOAST namespace so BOAST related class and methods are prefixed
with \emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that
produces the Fortran (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
(Listing~\ref{lst:BOAST_algebra_C}) output. First we define two variables
\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third
variable named \emph{l} is a one dimensional local array of length $27$, it
is indexed in the range $-5$ to $21$. All those Variables are affected to Ruby
variables of the corresponding name.

On Line~4 we declare those three variables. Variable \emph{i} is then affected
the value 5. On Line~6 the \emph{j} Ruby variable is used to store the BOAST
expression \emph{$i+5$}. This variable will the be used transparently through
the rest of the program.

On Line~8 we use variable \emph{k} to index into array \emph{l} using the
bracket operator. Would the array be multidimensional, the index would be comma
separated, similar to Fortran notation.

On the last line, a call to the \emph{sin} function is made through the creation
of a FuncCall object.  The possibility to use this method is declared using the
\emph{register\_funccall} method.

      \paragraph{Control Structures}

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_control},
        numbers=right }
\begin{lstlisting}
BOAST::register_funccall("modulo")
i = BOAST::Int("i")
j = BOAST::Int("j")
BOAST::pr j === 0
BOAST::pr BOAST::For( i, 0, 100 ) {
  BOAST::pr BOAST::If( BOAST::modulo(i,7) == 0 ) {
    BOAST::pr j === j + 1
  }
}
\end{lstlisting}

\begin{minipage}[b]{0.47\linewidth}
\centering
\lstset{style=customf,
        caption={Fortran output.},
        label={lst:BOAST_control_FORTRAN} }

\begin{lstlisting}
j=0
do i=0, 100, 1
  if(modulo(i, 7)==0) then
    j=j+1
  end if
end do
\end{lstlisting}
\end{minipage}
\hspace{0.08\linewidth}
\begin{minipage}[b]{0.40\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_control_C} }

\begin{lstlisting}
j=0;
for(i=0; i<=100; i+=1){
  if(modulo(i, 7)==0){
    j=j+1;
  }
}
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for control structures.}
\label{fig:BOAST_control}
\end{figure}

The classical control structures are implemented. \emph{If}, \emph{For},
\emph{While}, \emph{Case} are abstractions in BOAST matching the behavior of
corresponding control structures in other languages. An exception is the
\emph{For} in BOAST matches more closely the \emph{For} in Fortran than the one
in C.

Figure~\ref{fig:BOAST_control} shows some basic usages of the control
structures.  The example shows a C macro or function that behaves similarly to
the Fortran modulo intrinsic. The sample script (inefficiently) computes and
stores in \emph{j} the number of multiples of $7$ in the 0 to 100 range. It
uses the For and If control structures. A Ruby block is passed to each of those
constructs. This block is evaluated at the time the construct is printed. If
several such constructs are needed (in an \emph{if elsif else} case, for
instance) they can be explicitly passed as parameters using the \emph{lambda}
Ruby keyword.

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_procedure},
        numbers=right }
\begin{lstlisting}
n = Int("n", :dir=>:in)
ndat = Int("ndat", :dir=>:in)
x = Real("x", :dir=>:in, :dim=>[Dim(0, n-1), Dim(ndat)])
y = Real("y", :dir=>:out, :dim=>[Dim(ndat), Dim(0, n-1)])
p = Procedure("magicfilter", [n, ndat, x, y])
opn p
close p
\end{lstlisting}

\lstset{style=customf,
        caption={Fortran output.},
        label={lst:BOAST_procedure_FORTRAN} }

\begin{lstlisting}
SUBROUTINE magicfilter(n, ndat, x, y)
  integer(kind=4), intent(in) :: n
  integer(kind=4), intent(in) :: ndat
  real(kind=8), intent(in), dimension(0:n-(1), ndat) :: x
  real(kind=8), intent(out), dimension(ndat, 0:n-(1)) :: y
END SUBROUTINE magicfilter
\end{lstlisting}
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_procedure_C} }

\begin{lstlisting}
void magicfilter(const int32_t n, const int32_t ndat,
                 const double * x, double * y) {
}
\end{lstlisting}
\caption{BOAST Code Snippet for Procedure.}
\label{fig:BOAST_procedure}
\end{figure}

The last control structure is Procedure. It describes either procedures or
functions. Code in Figure~\ref{fig:BOAST_procedure} presents the use of this
abstraction. It illustrates the signature of a real kernel from BigDFT. This
kernel uses an input array \emph{x} and an output array \emph{y}, both composed
of double precision numbers. Those arrays have two dimensions which depend on
input variables \emph{n} and \emph{ndat}. We can see here the use of the
\emph{opn} and \emph{close} keywords that are used to print a control structure
without an associated Ruby block. This time we placed ourselves inside the
BOAST namespace.

The generated outputs in Fortran (Listing~\ref{lst:BOAST_procedure_FORTRAN})
and C (Listing~\ref{lst:BOAST_procedure_C}) show the difference in
meta-information that is kept between both versions.

%%% BOAST RUN-TIME %%%
\subsection{BOAST Run-time}

In the previous Section we presented the BOAST's language. This allows us to
describe procedures and functions and to meta-program them using Ruby.  Each
version has to be compiled, linked, and executed to assess its performance in
order to find the best version of a computing kernel.  This can be very time
consuming if this process cannot be automated. By enabling more versions for
evaluation, the automation will bring improved portability, better performance,
and in the end will improve the productivity of the developer.

In this section we will present the different aspects of BOAST run-time that
allow this automation. Those aspects are multi-target language generation
(\emph{performance}, \emph{portability}), kernel compilation
(\emph{productivity}, \emph{performance}), kernel execution
(\emph{productivity}, \emph{performance}) and last, kernel tracer, dumper and
replay for non regression tests (\emph{productivity}).

%%% MULTI-TARGET LANGUAGE GENERATION %%%
\subsubsection{Multi-target Language Generation\label{sec:multitarget}}
Language availability and performance varies between platforms. It is thus
important to express computing kernels in different languages, based on the
availability and their respective merits on the target platform. Some languages
have additional features, such as languages that target GPUs (OpenCL, CUDA), or
languages that support \FD{BV}{multi-threading?} threading paradigms (OpenMP). The developer should be
given tools to determine what kind of language is currently used in order to be
able to use those additional features.

Similarly the target language must be changed with ease in order to compare
different alternatives. Two methods are dedicated to this task: \emph{set\_lang}
and \emph{get\_lang}. The target language can also be set through an environment
variable before launching BOAST, allowing for easy command line scripting.

%%% COMPILATION %%%
\subsubsection{Compilation}
Compilation of the generated kernels must also be very flexible because HPC
application developers may encounter platforms with very diverse compilation
environments.  Proprietary and dedicated compilers are common on HPC
infrastructures. Thus BOAST build system exhibits similar behavior to common
build systems. Compilers and their compile/build options can be specified at
several places. The list by increasing order of precedence includes: BOAST
configuration file, environment variables, and at kernel build time.

This way the framework to test different compiler optimizations is completely
available and performance study can include both kernel related parameters and
compilation related parameters. This behavior is contained in the
\emph{CKernel} class of BOAST. When instantiating this class, a BOAST
\emph{Procedure} representing the entry point of the kernel is specified.

%%% EXECUTION %%%
\subsubsection{Execution}

The next logical step is to benchmark the built kernel. BOAST offers a simple
way to run a kernel that was successfully built without the need to fork
another process. A built BOAST \emph{CKernel} exposes a run method that
accepts arguments corresponding to the BOAST \emph{Procedure} used to
instantiate the kernel. Arrays must be instances of \emph{NArray} which are
numerical arrays that use C arrays underneath.

Arrays which correspond to output parameters will be modified during the
execution of the kernel so results can be checked. This allows for easy non
regression testing. The run method also returns information (and result for
kernels that are functions as well as output scalars) about the run. For
instance, one such information includes the run time of the kernel and it is
obtained using the system-wide real-time clock. Other probes can be inserted at
compile-time, if needed.  BOAST also supports PAPI~\cite{mucci1999papi} to
capture hardware performance counters during each kernel execution.

%%% KERNEL REPLAY %%%
\subsubsection{Kernel Replay}

The non regression methodology presented before is viable provided input data
can be generated at run-time. For instance, in the case of the Laplace kernel,
generating an input image and the corresponding reference output image, using a
reference implementation, is easy. Unfortunately it is not always possible,
because some applications have complex data patterns that are difficult to
synthesize without running the full application. Thus BOAST offers a way to load
binary data from the file system and use them as inputs of a kernel. Outputs can
also be checked against those binary data thus enabling (almost) data oblivious
non regression testing.

In order to use this methodology one has to be able to trace an application to
get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
in the next subsection.

%%% NON REGRESSION TESTING USING TRACE DEBUGGING %%%
\subsection{Non Regression Testing Using Trace Debugging\abel{sec:gputrace}}

Debugging applications running on GPU environments is well-recognized as a hard
and time-consuming activity. In complement with BOAST, we designed a trace-based
debugging tool that simplifies this porting operation. The tool relies on BOAST
support of multi-target code generation (Section~\ref{sec:multitarget}), used to
validate an application from one GPU-programming framework to another or between
different implementations using the same programming model. A case-study of the
port of a CUDA application to OpenCL is presented in
Section~\ref{subsec:specfem}.

The idea behind this tool is based on the assumption that the different GPU
ports of the code should perform the very same operations, at least at the
logical level (the APIs will have \emph{implementation} differences, but they
should offer nonetheless same functionalities). The usage of BOAST framework
sustains this assumption, as both sets of kernels should be generated from the
same source code.

Hence, the verification and validation of the new port can be narrowed down to
asserting that both codes apply the same operations on the GPU. And the
debugging part will consist in understanding what diverges. To that purpose, we
developed \code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
\code{GPUTrace} dynamically preloads a library between the application and the
GPU library, and collects the function name, execution range and argument
values (input \emph{and} output) of the relevant function calls. These
information are traced in a unified format for all the APIs. This is a custom
trace format. Listing~\ref{lst:GPUTrace} presents a sample output generated
during SPECFEM3D tracing.

However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
to be \emph{state-full}. Indeed, most of API parameters are handles to opaque
types. So, in order to generate meaningful traces, \code{GPUTrace} gathers
information about these objects at creation time (handle value, buffer creation
size and attributes, kernel name and prototype, \etc{}) and re-injects this
information when the objects used. A state-less implementation of
\code{GPUTrace} would highly rely on the GPU libraries introspection
capabilities, which seem not possible at the moment.

\lstset{style=customgputrace,
        caption={GPUTrace sample trace}, 
        label={lst:GPUTrace},
        numbers=left}
\begin{lstlisting}
New kernel: update_potential_kernel
...
New buffer #94, 2Mb, READ_WRITE (0x74d0420)
New buffer #95, 2Mb, READ_WRITE (0x74d08d0)
New buffer #96, 2Mb, READ_WRITE (0x74d0d80)
...
Buffer #94 written, 2314764b at +0b 
Buffer #95 written, 2314764b at +0b
Buffer #96 written, 2314764b at +0b
...
update_disp_veloc_kernel<128,1><4522,1>(
     float *displ=<buffer #94 1.00e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
     float *veloc=<buffer #95 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
     float *accel=<buffer #96 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
     const int size=<578691>
     const float deltat=<1.365914e-04>
     const float deltatsqover2=<9.328606e-09>
     const float deltatover2=<6.829570e-05>
     ----
     <out> float *displ=<buffer #94 1.000e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
     <out> float *veloc=<buffer #95 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
     <out> float *accel=<buffer #96 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
);
\end{lstlisting}

Once two call traces have been generated by \code{GPUTrace}, the user can
compare them with a graphical \code{diff} tool, and spot the different porting
mistakes: two parameters reversed, an offset incorrectly applied, \etc{}

By default, \code{GPUTrace} only prints a unique identifier (the creation index)
for the memory buffers.  Additionally, \code{GPUTrace} supports several modifier
flags. One flag can be activated to append the first bits of the buffer to the
trace, printed in the right format, for visual inspection. Another flag can be
set to dump the whole content of the buffer into a file, for a full inspection.

This last option is also useful to generate \emph{replay buffers} for BOAST
kernel execution. With a set of filters based on kernel names and execution
counters, developers can precisely select which kernel execution parameters
should be dumped, for further reuse as real-case benchmarks and non-regression
testing.

%%% BOAST USE CASES %%%
\section{BOAST Use Cases\label{use_cases}}

In this section we will present the benefits of using BOAST on the Laplace
motivating example as well as on two scientific applications. The first one,
BigDFT~\cite{genovese2008daubechies}, uses BOAST in order to develop new
functionalities with performance portability in mind. The second one,
SPECFEM3D~\cite{komatitsch2011fluid}, uses BOAST to factorize OpenCL and CUDA
development, while having robust non regression tests.

%%% LAPLACE FILTER KERNEL %%%
\subsection{Laplace Filter Kernel}

Section~\ref{sec:laplace} presented the Laplace motivating example. From this
section we know that a number of optimizations can have an impact on the
performance of the kernel on the Mali architecture. But, what is the impact in
other architectures? And, are there additional optimizations that can impact
the performance?

%%% OPTIMIZATION SPACE %%%
\subsubsection{Optimization Space}

The list of already identified optimizations are vectorization, intermediary
data type, number of pixels processed, and synthesizing loads.  To create our
generic implementation we decided to work at the component level rather than the
pixel level. This approach leads to more flexibility and genericity when
applying optimizations. We also decided to study the impact of another parameter
which is the number of components to process on the column direction. This leads
to being able to process tiles instead of only rows.

Here are the parameters we finally selected for our kernel optimization and
their possible values:
\begin{itemize}
\item \emph{x\_component\_number}: a positive integer
\item \emph{y\_component\_number}: a positive integer
\item \emph{vector\_length}: 1, 2, 4, 8 or 16
\item \emph{temporary\_size}: 2 or 4
\item \emph{synthesize\_loads}: \emph{true} or \emph{false}
\item \emph{vector\_recompute}: \emph{true} or \emph{false}
\end{itemize}
The last parameter is used when \emph{x\_component\_number} is not divisible by
\emph{vector\_length}. Two solutions are possible then, divide the remainder
of the division in vectors of smaller sizes (\emph{vector\_recompute} =
\emph{false}) or load more data and compute useless values in vectors of the
specified size (\emph{vector\_recompute} = \emph{true}).
This last option mimics the behavior of the ARM implementation, although when
working at the component level it may not be a valuable thing to try.

    \subsubsection{Performance Results}

Table~\ref{tbl:ARM} show the best results obtained by ARM on different images
compared to the naive implementation and to the best version version BOAST
found. It shows that the generated version systematically outperforms the hand
optimized version. As far as the optimization options are concerned, the results
are disappointing: the same kernel configuration is the best for all image
sizes. This kernel uses \emph{x\_component\_number} = 16,
\emph{y\_component\_number} = 1, \emph{vector\_length} = 16,
\emph{temporary\_size} = 2 and \emph{synthesize\_loads} = false.
\emph{vector\_recompute} because \emph{x\_component\_number}
== \emph{vector\_length}. Those results show that, when working on full
vectors, synthesizing the loads is harmful to performance and the programmer is
better of trusting the cache to load each vector in one cycle without
compromising the bandwidth. The results shown here are the best of four runs for
each configuration.

\begin{table}
\caption{Best performance of ARM Laplace kernel.}
\label{tbl:ARM}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size  & Naive (s) & Best (s) & Accel. & BOAST (s) & Accel. \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
% \hline
2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
% \hline
2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
% \hline
5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
% \hline
7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
% \hline
\end{tabular}
\end{table}

But what if we run our benchmark on other architectures? Table~\ref{tbl:other}
shows the results obtained when running our BOAST implementation on other
architectures. The chosen architectures include an Intel i7-2760QM CPU (Sandy
Bridge architecture) that supports OpenCL 1.2 and a system with an NVIDIA
gtx680 GPU that supports OpenCL 1.1. We see that the performance ratio between
the different architectures is stable across image sizes.

The optimization parameters selected are not the same for those architectures.
Indeed, the Intel CPU favors kernels that have the parameters,
\emph{x\_compo-nent\_number} = 16, \emph{vector\_length} = 8,
\emph{temporary\_size} = 2 and \emph{synthesize\_loads} = false. Once again
\emph{vector\_recompute} does not apply. \emph{y\_component\_number} varies
from 4 to 2 when image size increase, thus decreasing task granularity as the
global work size increase. This result is unexpected and understanding why this
happen could be interesting. The NVIDIA GPU favors processing square tiles:
\emph{x\_component\_number} = 4, \emph{y\_component\_number} = 4,
\emph{vector\_length} = 4, \emph{temporary\_size} = 2 and
\emph{synthesize\_loads} = false. Once more, \emph{vector\_recompute} has no
meaning.

\begin{table}
\caption{Best performance of Laplace Kernel on several architectures.}
\label{tbl:other}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size & ARM & Intel & Ratio & NV & Ratio \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
% \hline
2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
% \hline
2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
% \hline
5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
% \hline
7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
% \hline
\end{tabular}
\end{table}

    \subsubsection{Laplace Conclusion}

In this Subsection we have shown the interest of BOAST in optimizing a
well-known algorithm across different architectures. Chosen optimization
combinations are highly dependent on the target architecture. But BOAST
simplifies and speeds up the identification of the correct combination at the
cost of metaprogramming. In the next Subsections we will show that our
methodology also applies to real applications.

  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using
BOAST}

In 2005, the EU FP6-STREP-NEST BigDFT~\cite{genovese2008daubechies} project
funded a consortium of four European laboratories (L\_Sim - CEA Grenoble, Basel
University - Switzerland, Louvain-la-Neuve University - Belgium and Kiel
University - Germany), with the aim of developing a novel approach for DFT
calculations based on Daubechies wavelets.  Rather than simply building a DFT
code from scratch, the objective of this three-years project was to test the
potential benefit of a new formalism in the context of electronic structure
calculations.

As a matter of fact, Daubechies wavelets exhibit a set of properties which make
them ideal for a precise and optimized DFT approach. In particular, their
systematicity allows to provide a reliable basis set for high-precision results,
whereas their locality (both in real and reciprocal space) is highly desired to
improve the efficiency and the flexibility of the processing. Indeed, a
localized basis set allows to optimize the number of degrees of freedom for a
required accuracy~\cite{genovese2008daubechies}, which is highly desirable given
the complexity and inhomogeneity of the systems under investigation nowadays.

Despite that the application is mainly written in Fortran, it currently includes
360 kLOC and 70 kLOC of Fortran and C languages, respectively, accounting for
more than 50\% of the code base.  It is a parallel application based on the
standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports
CUDA~\cite{cuda} and OpenCL~\cite{opencl}.  In the recent years this code has
been used for many scientific applications, and its development and user
consortium is continuously growing.  Massively parallel computations are
routinely executed with the BigDFT code, either in homogeneous or hybrid
architectures.  In 2009, the French Bull-Fourier award was attributed for the
implementation of the hybrid version of BigDFT~\cite{genovese2009density}.

%As it is used on systems that may
%have very different architectures, it is of great importance to be able to
%optimize and run the application according to the specific underlying platform.

    \subsubsection{BigDFT}

In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to
wavefunctions (orbitals), which are represented by arrays of floating point
numbers. In wavelets formalism, the operators are written via convolutions with
short, separable filters.  The detailed description of how these operations are
defined is beyond the scope of this paper and can be found in the BigDFT
reference paper~\cite{genovese2008daubechies}. Convolutions are basic operations
of lots of scientific application codes, like for example finite differences
approaches, which are universally used in computational physics.

% Convolutions are among the basic processing blocks of BigDFT. Special care
% has to be taken regarding their performances. 
The CPU convolutions of BigDFT have thus been thoroughly optimized.  In a recent
paper~\cite{videau2013optimizing}, the optimization of the CPU convolutions of
BigDFT has been extensively considered.
%BigDFT is characterized by its extensive use of convolution
%operators~\cite{nussbaumer1982fast} applied on large arrays of data.  
One example of a specific convolution, called
MagicFilter~\cite{Genovese|CAS2010}, can be seen in
Listing~\ref{lst:conv_example}.  It applies a filter $filt$ to the data set
$in$ and then stores the result in the data set $out$ with a
transposition~\cite{Goedecker1993}.

\lstset{language=C, 
  basicstyle=\scriptsize, 
  backgroundcolor=\color{white}, 
  frame=single, 
  captionpos=b,
  caption={MagicFilter.}, 
  label={lst:conv_example},
  numbers=right}

\begin{lstlisting}
double filt[16] = {F0, F1, ... , F15};
void magicfilter(int n, int ndat, double* in, double* out){
  double temp;
  int m;
  for( j = 0; j < ndat; j++) {
    for( i = 0; i < n; i++) {
      temp = 0;
      for( k = 0; k < 16; k++) {
        m = (i-7+k)%n
        temp += in[m + j*n] * filt[k];
      }
      out [j + i*ndat] = temp ;
    } 
  }
} 
\end{lstlisting}

As we can see, there are three nested loops working on arrays whose sizes vary.
Various optimizations can be applied to this treatment and may focus on the
loop structure, as well as on the size of the data. 
%In this paper we focus on this particular convolution and explore different
%optimizations whose nature is presented in detail in the next section.

    \subsubsection{A Generic Convolution Library}

The number of convolution kernels needed in BigDFT has been continuously
growing in the recent years.  Various boundary conditions and functionalities
have been added, making the BigDFT more and more powerful in terms of
scientific applications.  However, the cost of maintenance and development of
the convolutions is always a delicate point to be considered while including a
new functionality.  The convolution patterns are usually rather similar,
leading to code duplication and difficulties in code maintainability.

Therefore it appears very interesting to benefit from an automatic tool to drive
the implementation and the generation of new convolutions.  This would lead to
an optimized code, adapted to different computing platforms, that is optimally
factorized.  In addition to this point, the help of such code generator is also
important to build \emph{new science}: the cost of implementing new
convolutions would become so little that other functionalities (for example,
the generalization of the BigDFT convolutions to Neumann boundary conditions
or the usage of wavelet-on-the-interval basis) can be added with limited
manpower.

For these reasons, a convolution library has been engineered with the help of
BOAST. The detailed API of the library is beyond the scope of this article. The
spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
and called from high-level operations. The basic blocks will be comprised of
unidimensional wavelet transforms and convolutions applied to multidimensional
arrays. Combining those blocks will yield multidimensional transforms.

Each of the building blocks of these convolution libraries is optimally tuned by
BOAST by choosing the sources providing the optimal kernel for the chosen
computing platform.  The sources of these kernels are then collected and
compiled to meet the API of the library. A library written in this way might
have an impact going largely beyond the community of BigDFT developers.

The interest in having a robust and optimally tuned library in this scientific
field is therefore evident. Techniques are under investigation to also provide
end users with fine-tuned binaries rather than the source codes, such that more
aggressive inter procedural optimization can be performed. Indeed, BOAST finds
the optimal source code for a given kernel and compiler configuration but one
could imagine using binary optimizer to the compiled binaries. Those optimizers
could be coupled to BOAST and the output of the process would be the final
binary rather than the source code. This experience therefore would be a first
step toward the release of a tunable optimized convolution library oriented to
computational physics communities.

    \subsubsection{Performance Report}

Several kernels have already been implemented in BOAST for the convolution
library. Figure~\ref{fig:synthesis} shows the performance of the wavelet
transform operation as a factor of the unrolling length of the outer loop, the
language used to implement it as well as the activation or not of the OpenMP
parallelization. Results are given as a speedup compared to the sequential hand
tuned implementation that can be found in BigDFT. These tests were run on the
Intel Xeon X5550 that was used to hand optimize the code.

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{Res_synthesis}
\end{center}
\caption{Impact of Unrolling, Language Used and OpenMP on a Wavelet Transform Code.}
\label{fig:synthesis}
\end{figure}

We can see that this function is better optimized using Fortran and small
unrolling factors. In the hand optimized version the unrolling factor was chosen
much too high (a factor of 12 was used). This factor might have been optimal at
the time the procedure was optimized (compiler version changed in the meantime) but
since then the environment changed. Other optimizations have also been
incorporated in the BOAST sources, like the systematic inner loop unrolling, and
those could also help increase performance while limiting the interest of the
outer loop unrolling.

Nonetheless what is interesting from the physicist point of view is that the
generated source will give its better performance on a whole range of
architectures/compilers combinations than that of the hand tuned code.

%   BigDFT implementation of convolution/wavelets is a reference but is limited
%   in scope (1 wavelet family ands its derived operators).  Generalizing to
%   other wavelet families long and difficult if performance are to remain
%   good.
% 
% Idea: \begin{itemize} \item write a generic convolution library in BOAST
% \item BLAS like interface \item Provide a wide variety of building blocks
% \item Optimize thos blocks using BOAST \item Build either from source or
% directly link with binary \item Release complete library \end{itemize}
% 
% 
%   Characteristic use case driven
% 
%   Convolution Library: first step toward a dsl dedicated to wavelet
%   processing for Physics computing.

  \subsection{Porting SPECFEM3D to OpenCL using BOAST}
  \label{subsec:specfem}

In this last subsection, we study a free seismic wave propagation simulator,
SPECFEM3D, in its GLOBE flavor\footnote{Specfem3d Globe --- CIG
\url{http://www.geodynamics.org/cig/software/specfem3d-globe}}. SPECFEM3D
simulates seismic wave propagation at the local or regional scale based upon
spectral-element method (SEM), with very good accuracy and convergence
properties. It is a reference application for supercomputer benchmarking,
thanks to its good scaling capabilities.

    \subsubsection{SPECFEM3D}

When we started to work on the project (version v2.1 of July 2013) it
supported graphics card GPU acceleration through NVidia CUDA. This
GPU support came in addition to the MPI support implemented to enable
multi-CPU parallel computing. Most of SPECFEM3D code-base is written in
Fortran 2003 and only the GPU related parts are written in
C.  The split between CPU and GPU code was done at a rather fine
grain, as the application counted more than 40 GPU kernels. Some of
them were quite simple (\eg performing a few vector operations, but at
massively parallel scale), while at the other side of the spectrum,
some complex kernels took more than 80 parameters and perform very
specific physical transformations.

Because of the complexity of wave propagation (and of the application
architecture), it is hard to impossible to define unit tests. Hence,
the application is validated by the accuracy of the results it
produces, that is, the accuracy of seismograms of the simulated
earthquakes, in comparison with the actual ones. Non-regression
testing (after new developments) is also based on these seismograms,
with measurements of the relative error between two identical
simulations.

As part of the Mont-Blanc project, we had to port SPECFEM3D to OpenCL, so
that it could be used to benchmark Mont-Blanc HPC platforms.

\subsubsection{Porting to OpenCL}

\paragraph{Porting kernels to BOAST}

Nvidia CUDA and OpenCL are based on the same programming model: a
massively parallel accelerator running in disjoint, non addressable
memory environment. Thanks to that proximity, we have been able to carry
out most of the porting task with only a limited knowledge about
SPECFEM3D internal physics.

This lack of SPECFEM3D internal knowledge led us to be particularly
careful to the path we undertook for the port, as we would have
been unable to understand how and why the application was not
operating properly, if it was to fail.

Hence, our first milestone in the port was the translation of
SPECFEM3D's CUDA kernels into BOAST EDSL. This way, we could ask BOAST
framework to generate a CUDA version of the kernels, plug them back
into SPECFEM3D and get (after fixing compile-time errors---prototypes
and naming mistakes mainly) a first set of SPECFEM3D seismograms.

As we had expected, the seismograms were erroneous. But with the help
of shell scripts and BOAST framework ability to store and provide the
kernels' original source code, we built a set of SPECFEM3D binaries
including only \emph{one} BOAST-generated kernel, with all others
reference-kernels. Running and validating all these binaries enabled
use to pinpoint the misbehaving kernels. We finished the debugging
with a side-by-side comparison that highlighted the coding mistakes.

\paragraph{Porting run-time to OpenCL} The second part of the port
consisted in the translation of the CPU-side of the application, from
CUDA API to OpenCL API. Most of the functions of the interfaces are
very similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of OpenCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations, and 270 releases.

Once the transformations were applied,
compilation errors fixed, and OpenCL unsuccessful function calls solved,
the application managed to complete its execution and generate a first
set of seismograms. And again, as expected and feared, these
seismograms were not valid as their shape was completely different from the
reference ones.

As we had already validated BOAST-generated kernels (and trusted CUDA and OpenCL
versions to be semantically identical), we knew that the bugs were now in the
usage of the run-time, and we had to find a way to understand where SPECFEM3D's
CUDA version of the code diverged from its OpenCL counterpart. To help us in
that purpose, we had a strong assumption: both versions of the code were
supposed to perform exactly the same operations, with the same ``logical''
parameters (the APIs have \emph{implementation} differences, for instance OpenCL
has two memory transfer functions, \code{clEnqueueReadBuffer},
\code{clEnqueueWriteBuffer}, whereas CUDA has only one, with a direction
parameter \code{cudaMemcpy(..., dir)}, but above that, it is the same
functionality).

Hence, our idea for locating the execution problems was to make sure
that both execution actually did the same thing. As the OpenCL results
were invalid, we knew the executions would diverge at one or several
points.

\paragraph{Debugging OpenCL Execution: \code{GPUTrace}}

With the help of \code{GPUTrace} (described in Section~\ref{sec:gputrace}), we
could confront CUDA and OpenCL execution traces with a graphical \code{diff}
tool, and spot the different porting mistakes: some parameters reversed,
offsets incorrectly applied, \etc{}.

One last problem remained, clearly highlighted by the seismograms not
matching perfectly (they had a similar shape, but with a reduced
intensity). We added more verbosity to \code{GPUTrace} output: first
the initial bits of the GPU memory buffers, then their full
content. The drift was visible in the trace, but it was nonetheless
unclear where it started. We finally got it after hours of code review
of BOAST kernels and OpenCL code. One kernel was
\emph{three}-dimensional, whereas the others were two-dimensional. But
for all of them, only two dimensions were passed, and one was
missing.
%As it was the same person who coded the OpenCL port and the GPU
%tracer, this third dimension was forgotten in both codes.

\paragraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
merged in SPECFEM3D's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and N Intel
Xeon processors, we measured similar results between the original
CUDA version and our BOAST-CUDA version. With the same set of
optimization flags, BOAST CUDA and OpenCL version reported similar
execution time spans. The best execution speed was achieved with CUDA
version though (25\% higher than OpenCL), as one optimization parameter
(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to OpenCL run-time, as of version
1.1. This parameter, in addition to specifying the work group size (which can
be done in OpenCL), also constrains the number of work group that must run in
parallel on a multiprocessor. This value is set to 7 in CUDA. This means that
the compiler must be very conservative on register usage in order to allow this
parallelism which allows better overlapping of communications end computations.
This functionality is not supported in OpenCL.

By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, partly
because of code duplication, also removing manually unrolled
loops). This is beneficial for SPECFEM3D as it improves the readability
and maintainability of its source code.

We have also been able to enhance SPECFEM3D's non-regression test-suite
by adding per-kernel non-regression tests. This was done with the help
of \code{GPUTrace}, that we used to capture all the input parameters
of a particular \emph{valid} kernel execution, as well as the output
values. Then, during the non-regression testing, BOAST framework
loads these buffer files, allocates GPU memory and initializes it through
CUDA or OpenCL run-time, and triggers the kernel execution. A comparison
of the output values (for instance against a maximal error level)
validates the non-regression.

In the same mindset, we provided SPECFEM3D test-suite with kernel
performance evaluation mechanisms. These tests will help developers to
try new optimizations in kernel code and measure their impact, without
executing the whole application.

\section{Related Work\label{sec:related}}

Code generation and auto-tuning techniques are not new. Nonetheless, with recent
developments in hardware, and the HPC landscape being as diverse as it is now,
there is a renewed interest in the field. This related work Section is split in
three parts, focusing first on Auto-tuning frameworks. Tools that provide a DSL
to describe computing kernels will then be presented. Last, optimization space
pruners and their ties to auto-tuning will be introduced.

  \subsection{Application Auto-Tuning}

  The most convenient way to obtain an application that can be auto-tuned on a
given platform is to base this application on a widely used computing library.
BLAS~\cite{dongarra1990set} and LAPACK~\cite{laug} are such libraries. These
libraries are either hand tuned for selected platforms or have auto-tuned
implementations. Atlas~\cite{whaley04} is an auto-tuned implementation of
BLAS/LAPACK. ATLAS authors defined the \emph{Automated Empirical Optimization of
Software} methodology that we implemented with BOAST. Their kernel generation is
done using macro-functions in C.

  Nonetheless many application formalisms cannot be reduced to standardized
library or border on what could be considered edge cases for those libraries and
not as optimized as more general cases. Orio~\cite{Hart2009:Orio} is an
auto-tuning framework that has an approach close of that of BOAST but they are
based on an annotated DSL describing loop transformation rather than a more
generic EDSL. Halide~\cite{ragan2013halide} is an auto-tuning framework
dedicated to image processing. It can also be used to describe other
operations on memory buffers. Those two frameworks propose automated search
space exploration to find the best version of a kernel.
LGen~\cite{spampinato2014basic} is a compiler that generates linear algebra
programs for small fixed size problems. Knowing the problem size it fully
unrolls and vectorizes loops, yielding better performance than state of the
art generic implementations.

  \subsection{Kernel Description DSL}

  The idea to describe computing kernels using a DSL has been already explored.
SPIRAL~\cite{puschel2004spiral} is a decade old generation
framework for signal processing. It uses a proprietary DSL called SPL (Signal
Processing Language), to describe a DSP algorithm. This DSL is then transformed
into efficient programs in high level languages like C or Fortran.
POET~\cite{yi2007poet} also uses a DSL to describe custom code transformations,
like loop unrolling, loop blocking and loop interchange. Those transformations
can be parametrized in order to tune the application. Orio~\cite{Hart2009:Orio}
can be compared to POET as it aims at describing possible code transformations
using a DSL. All those approaches are very different from ours as they put the
emphasis on compilation techniques whereas BOAST relies on the user to express
the different optimizations.

  Halide~\cite{ragan2013halide} is closer in some ways to our approach as it
uses an embedded C++ DSL to describe image processing algorithm. This DSL allows
decoupling the algorithm description from its scheduling. Each pixel in the
resulting image has a completely defined dependency tree with regard to pixels
in the input image (and intermediary results). During generation, depending on
memory and computing cost, some values are recomputed rather than fetched from
memory. We used Halide to implement the \emph{magicfilter} of BigDFT, but
unfortunately results were 4 to 5 times slower than the one we obtained with
BOAST. We speculate that the three-dimensional nature of our convolutions, as
well as the filter length, can be considered edge cases in Halide and quite far
from the intended target.

  Heterogeneous Programming Library~\cite{vinas2013exploiting} (HPL) is maybe
the most similar to BOAST. It is an autotuning kernel library in C++ that targets
OpenCL. It uses an EDSL in C++ to describe generic computing kernels. Like when
using BOAST, optimization are left to the users and written directly using the
EDSL. The syntax is very close to real C++ with the classical control
structures having a trailing underscore. But whereas BOAST re-interprets the
same code while changing the environment, in HPL each expression is stored in a
tree and the tree is reevaluated in a different environment.

  \subsection{Optimization Space Pruner}

  Once auto-tuning techniques are used, the parameter space explodes and the
systematic sampling rapidly becomes impossible to achieve. The cost of finding
the optimal kernel parameters and environment parameters (compiler flags, used
language, ...) is rapidly prohibitive. Dedicated frameworks have been developed
to address this problem. Adaptive Sampling Kit (ASK)~\cite{castro2013adaptive}
is one such tool. It reduces the number of samples needed by creating a model of
the performance and by minimizing the number of samples needed to find the
parameters of this model. Several models and sampling techniques are
implemented.

  Collective Mind~\cite{fursin:hal-01054763} proposes similar techniques to
solve the problem but also stores and the results and complete experimental
setups in databases for future reference and reproducibility. This database
approach also enables easy parallelization of the experimental process.
Collective Mind also proposes collections of flags for many available
compilers/versions easing the exploration process.

\section{Conclusion and Future Works}

In this paper we presented the BOAST infrastructure as well as its application
to three use cases from the Mont-Blanc project. Results are encouraging as
BOAST proved to be a powerful and flexible tool that allowed gains in
performance compared to hand tuned codes. Performance portability of those
codes is also improved.

Future development will focus on three goals. Interfacing with binary analysis
tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop
to guide optimization. Interfacing with search space modellers/pruners in order
to optimize the search of the optimal version of a kernel. Work should also
continue on improving the support for vector codes. For instance producing a
collection of small to medium useful vector patterns (transposition, for
instance) in BOAST could really help users develop vectorized version of their
algorithm.

\section*{Acknowledgments}

The research leading to these results has received funding from the European
Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement
\Num 288777 and 610402.

%\section*{References}

\bibliography{BOAST}

\end{document}