### 2017-04-25 Paper2/answer-reviewers.tex
ajout

\documentclass[10pt]{article}

\usepackage{fullpage}
\usepackage{url}

\begin{document}

\noindent
{\Large\bf ANSWERS TO THE REVIEWERS}\\~\\

We thank the  reviewers for their interest in our work and for their helpful comments that will greatly improve  the manuscript. We have  tried  to do  our  best to  respond to  the  points  raised. The reviewers have brought  up  some  good  points  and  we  appreciate  the  opportunity  to  clarify  our research objectives and results. As  indicated  below,  we  have  checked  all  the  general  and  specific  comments  provided  by  the reviewers and have made necessary changes accordingly to their indications.\\

\noindent{\large\bf Anonymous reviewers \#1}\\~\\

\end{document}
%%%