### 2017-08-30 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni: kernel update

\documentclass[10pt,twoside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper, bottom=2.5cm, top=2cm, left=2cm, right=2cm}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx,subfigure, array}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref,url}
\usepackage[hang,  bf, margin=19pt, tableposition=top,textfont=it]{caption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multicol,multirow}
\usepackage{subfigure}
\usepackage{fancyvrb,newverbs,xcolor}

% COMMANDS
\newcommand{\N}{\mathbb{N}} 
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{property}{Property}
\newtheorem{proof}{Proof}

\newlength{\textlarg}
\newcommand{\barre}[1]{%
   \settowidth{\textlarg}{#1}
   #1\hspace{-\textlarg}\rule[0.5ex]{\textlarg}{0.5pt}}

\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{break}}
\algnewcommand\Break[1]{\State \algorithmicbreak(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

% COLORS
\definecolor{bleugris}{RGB}{43,78,103}
\definecolor{codegreen}{rgb}{0.56,0.74,0.56}
\definecolor{backcolor}{rgb}{0.95,0.95,0.96}
\definecolor{lightgray}{gray}{0.35}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codered}{rgb}{0.87,0.19,0.39}
\definecolor{cverbbg}{gray}{0.93}

\newenvironment{cverbatim}
 {\SaveVerbatim{cverb}}
 {\endSaveVerbatim
  \flushleft\fboxrule=0pt\fboxsep=.5em
  \colorbox{cverbbg}{\BUseVerbatim{cverb}}%
  \endflushleft
}
\newenvironment{lcverbatim}
 {\SaveVerbatim{cverb}}
 {\endSaveVerbatim
  \flushleft\fboxrule=0pt\fboxsep=.5em
  \colorbox{cverbbg}{%
    \makebox[\dimexpr\linewidth-2\fboxsep][l]{\BUseVerbatim{cverb}}%
  }
  \endflushleft
}

\newcommand{\ctexttt}[1]{\colorbox{cverbbg}{\texttt{#1}}}
\newverbcommand{\cverb}
  {\setbox\verbbox\hbox\bgroup}
  {\egroup\colorbox{cverbbg}{\box\verbbox}}

\lstdefinestyle{mystyle}{
	%backgroundcolor=\color{backcolor},
	commentstyle=\color{codegreen},
	keywordstyle=\color{codered},
	numberstyle=\tiny\color{codegray},
	basicstyle=\scriptsize,
	breaklines=true,
	breakatwhitespace=false,
	numbers=left,
	tabsize=2
}

\lstset{style=mystyle}

\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox[cmyk]{0.43, 0.35, 0.35,0.01}{\parbox{\textwidth}{\hspace{15pt}#3}}}
\DeclareCaptionFormat{listing2}{\colorbox[cmyk]{0.9, 0.9, 0.9,0.01}{\parbox{\textwidth}{\hspace{15pt}#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white,singlelinecheck=false,margin=0pt,font={bf}}

\title{Hou10ni}
\date{}

\begin{document}
\maketitle

This document presents a study of hou10ni's kernel. We present results obtained with the performance evaluation tools TAU and Paje on the application.

\tableofcontents

\newpage

\section{Build and launch hou10ni}

Compilation:
\begin{cverbatim}
mkdir build && cd build
cmake ..
ccmake ..
make
cd .. 
\end{cverbatim}

\noindent Execution with 4 MPI processes:
\begin{cverbatim}
mpirun -np 4 build/hou10ni.out < HPC4E.dat
\end{cverbatim}

\section{How to use TAU on hou10ni}

\begin{cverbatim}
export TAU_MAKEFILE=$TAU_DIR/Makefile.
export TAU_OPTIONS='-optPdtF95Opts="-R free" -optVerbose'

cd build
cmake ..
ccmake ..
\end{cverbatim}

\noindent $\rightarrow$ Press "t" to change the compiler and chose \texttt{tau\_f90.sh}. Then you can tape \texttt{make} and launch the application.

\begin{cverbatim}
make
cd .. 
mpirun -np 4 tau_exec -T mpi  build/hou10ni.out < HPC4E.dat
\end{cverbatim}

\noindent To visualize the results with paraprof, 
\begin{cverbatim}
paraprof &
\end{cverbatim}
or
\begin{cverbatim}
paraprof --pack app.ppk
\end{cverbatim}
This last command will create the file \texttt{app.ppk}.

\section{Trace Visualisation with Jumpshot tool}

Jumpshot user guide: \url{ftp://ftp.mcs.anl.gov/pub/mpi/slog2/js4-usersguide.pdf}\\

\noindent To generate a trace of an application, follow the steps presented below.

\begin{cverbatim}
export TAU_TRACE=1
mpirun -np 4 tau_exec -T mpi  build/hou10ni.out < HPC4E.dat
tau_treemerge.pl
\end{cverbatim}

\noindent $\rightarrow$ Two files are created:  tau.trc and tau.edf

\begin{cverbatim}
tau2slog2 tau.trc tau.edf  -o app.slog2
jumpshot app.slog2
\end{cverbatim}

Figure \ref{fig:trace} shows the trace obtained with 4 MPI processes, TAU instrumentation and displayed with the Jumpshot tool. Figure \ref{fig:trace2} presents a zoom on the big portion and Figure \ref{fig:trace3} depicts the end of the trace. The purple/pink color represents the \texttt{MPI\_Wait} function. In Figure \ref{fig:trace3} we can see the different time spent in \texttt{MPI\_Wait} for all processes.

\begin{figure}[ht!]
\centering
\includegraphics[width=13cm]{IMAGES/jumpshot1}
\caption{Trace Visualisation with Jumpshot}
\label{fig:trace}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=13cm]{IMAGES/jumpshot2}
\caption{Zoom on the big portion in Figure \ref{fig:trace}}
\label{fig:trace2}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=15cm]{IMAGES/jumpshot3}
\caption{Zoom on the end of the trace}
\label{fig:trace3}
\end{figure}

%\begin{figure}[ht!]
%\centering
%\includegraphics[width=10cm]{IMAGES/jumpshot4}
%\caption{Zoom on the beginning of the trace}
%\label{fig:trace3}
%\end{figure}

\section{Profile Visualisation with paraprof}

It is possible to get a detailed summary in a text format. To record all information in a file \textit{profile\_summary.txt}, the command is
\begin{cverbatim}
pprof > profile_summary.txt
\end{cverbatim}

\noindent The file created looks like the following

{\scriptsize
\begin{verbatim}
Reading Profile files in profile.*

NODE 0;CONTEXT 0;THREAD 0:
---------------------------------------------------------------------------------------
%Time    Exclusive    Inclusive       #Call      #Subrs  Inclusive Name
              msec   total msec                          usec/call
---------------------------------------------------------------------------------------
 72.9            0     2:02.995       12300           0      10000 .TAU application => [CONTEXT] .TAU application
 72.9            0     2:02.995       12300           0      10000 [CONTEXT] .TAU application
 54.0     1:31.155     1:31.155        9116           0      10000 [SAMPLE] UNRESOLVED /lib64/libpthread-2.11.3.so
 54.0     1:31.122     1:31.122        9113           0       9999 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 23.3       39,314       39,314        1644           0      23914 MPI_Wait()
 23.3            0       39,300        3929           0      10003 MPI_Wait() => [CONTEXT] MPI_Wait()
 23.3            0       39,300        3929           0      10003 [CONTEXT] MPI_Wait()
 14.5       24,542       24,542        2454           0      10001 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 14.5       24,542       24,542        2454           0      10001 [SAMPLE] UNRESOLVED /home/esaillard/hou10ni2d/build/hou10ni.out
 10.7       18,132       18,132        1813           0      10002 [SAMPLE] MPID_nem_mpich_blocking_recv
 10.3       17,368       17,368        1737           0       9999 MPI_Wait() => [CONTEXT] MPI_Wait() => [SAMPLE] MPID_nem_mpich_blocking_recv
....

USER EVENTS Profile :NODE 0, CONTEXT 0, THREAD 0
---------------------------------------------------------------------------------------
NumSamples   MaxValue   MinValue  MeanValue  Std. Dev.  Event Name
---------------------------------------------------------------------------------------
         4          8          4          6          2  Message size for all-reduce
        76  1.375E+07          4  6.716E+05  2.489E+06  Message size for broadcast
      2022  5.426E+04          4  1.111E+04  1.929E+04  Message size received from all nodes
       822  5.426E+04       1024  2.732E+04  2.174E+04  Message size received in wait

....

NODE 1;CONTEXT 0;THREAD 0:
---------------------------------------------------------------------------------------
%Time    Exclusive    Inclusive       #Call      #Subrs  Inclusive Name
              msec   total msec                          usec/call
---------------------------------------------------------------------------------------
 93.7            0     2:38.072       15805           0      10001 .TAU application => [CONTEXT] .TAU application
 93.7            0     2:38.072       15805           0      10001 [CONTEXT] .TAU application
 61.9     1:44.423     1:44.423       10444           0       9998 [SAMPLE] UNRESOLVED /lib64/libpthread-2.11.3.so
 61.9     1:44.405     1:44.405       10443           0       9998 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 31.7       53,555       53,555        5351           0      10008 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 31.7       53,555       53,555        5351           0      10008 [SAMPLE] UNRESOLVED /home/esaillard/hou10ni2d/build/hou10ni.out
  4.2        7,074        7,074         424           0      16685 MPI_Recv()
  4.2            0        7,071         707           0      10002 MPI_Recv() => [CONTEXT] MPI_Recv()
  4.2            0        7,071         707           0      10002 [CONTEXT] MPI_Recv()
  3.0        5,135        5,135         516           0       9953 [SAMPLE] MPID_nem_mpich_blocking_recv
  2.3        3,807        3,807         380           0      10021 MPI_Recv() => [CONTEXT] MPI_Recv() => [SAMPLE] MPID_nem_mpich_blocking_recv
  1.5            0        2,550         255           0      10004 MPI_Bcast() => [CONTEXT] MPI_Bcast()
  1.5            0        2,550         255           0      10004 [CONTEXT] MPI_Bcast()
  1.5        2,543        2,543          76           0      33468 MPI_Bcast()

....
and so on
\end{verbatim}
}

\noindent Figures \ref{fig:1} and \ref{fig:2} show the entire application time with paraprof.

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot2}
\caption{Paraprof main window with stack bars together}
\label{fig:1}
\end{figure}

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot2bis}
\caption{Paraprof main window }
\label{fig:2}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{IMAGES/ScreenShotWait}
\caption{Time in one of MPI\_Wait function}
\label{fig:wait}
\end{figure}

\noindent {\bf{Statistics per process}} \\

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot5}
\caption{Statistics for processes 0 and 1}
\label{fig:5}
\end{figure}

\noindent {\bf{Comparison Window}} \\

To get a comparison window, click right on the text \textit{node 0} for example and chose \textit{"Add Thread to Comparison Window"}.

\begin{figure}[H]
\includegraphics[width=14cm]{IMAGES/ScreenShot3}
\caption{Comparison window}
\label{fig:3}
\end{figure}

\noindent {\bf{3D Visualization}} \\

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{IMAGES/ScreenShot4}
\caption{3D Visualization}
\label{fig:4}
\end{figure}

A complete manual to use TAU and paraprof is available at \url{https://www.cs.uoregon.edu/research/tau/tau-usersguide.pdf}.

\section{Trace Analysis with PAJE}

To use the tool, we first need to convert TAU traces to PAJE traces\footnote{For a tutorial on how to install tau2paje: https://github.com/schnorr/akypuera/wiki/TAUWithAkypuera}. After TAU generates tau.trc and tau.edf files, use tau2paje to generate tau.paje with the following command

\begin{cverbatim}
tau2paje --no-links --only-mpi --normalize-mpi tau.trc tau.edf > tau.paje
\end{cverbatim}

Then we use pj\_dump\footnote{For a tutorial on how to install pj dump: https://github.com/schnorr/pajeng/wiki/Install} to transform PAJE traces to CSV:

\begin{cverbatim}
pj_dump tau.paje | grep -e ^State -e ^Event > tau.csv
\end{cverbatim}

To check the head of the file:

\begin{cverbatim}
head tau.csv
\end{cverbatim}

\subsection*{Analysis of the CSV using R+dplyr+ggplot2}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :session *R* :exports both  
library(dplyr);
df <- read.csv("tau.csv", header=FALSE, strip.white=TRUE);
# get only lines starting with State
df <- df %>% filter(V1 == "State");
names(df) <- c("Type", "Thread", "Type2", "Start", "End", "Duration", "Imbric", "Region");
df <- df %>% select(-Type, -Type2, -Duration, -Imbric);
df$Thread <- as.integer(gsub("rank", "", as.character(df$Thread)));
head(df);
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
#+RESULTS:
:   Thread    Start      End        Region
: 1      3 0.013019 0.037004      MPI_Init
: 2      3 0.038628 0.038630 MPI_Comm_size
: 3      3 0.038637 0.038637 MPI_Comm_rank
: 4      3 0.041826 0.046533     MPI_Bcast
: 5      3 0.046534 0.046543     MPI_Bcast
: 6      3 0.046546 0.046550     MPI_Bcast
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :session *R* :exports both
library(ggplot2)
tstart = min(df$Start);
tend = max(df$End);

df %>%
  ggplot() + theme_bw(base_size=16) + xlab("Time [s]") +
  ylab("Thread") +
  theme(
     plot.margin = unit(c(0,0,0,0), "cm"),
     legend.margin = unit(.1, "line"),
     panel.grid = element_blank(),
     legend.position = "bottom",
     legend.title = element_blank()
   ) +
    coord_cartesian(xlim=c(tstart,tend)) +
    guides(fill = guide_legend(nrow = 2)) +
    geom_rect(alpha=1, aes(fill=Region,
                           xmin=Start,
                           xmax=End,
                           ymin=Thread,
                           ymax=Thread + 0.9))

tspan = tend - tstart;
tspan;
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
\end{lstlisting}
\vspace{-.5cm}

\begin{figure}[H]
\centering
\includegraphics[width=13cm]{IMAGES/tool1}
\caption{Trace}
\label{fig:5}
\end{figure}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :exports both :session *R* 
df %>%
  filter(Start >= tstart, End <= tend) %>%
  group_by(Thread, Region) %>%
  summarize(count = length(End), TDur = sum(End-Start)/tspan*100) %>%
  group_by(Thread) %>%
  summarize(t=sum(TDur)) %>%
  as.data.frame();
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
#+RESULTS:
:   Thread         t
: 1      0  3.221339
: 2      1 11.986336
: 3      2  6.962705
: 4      3 10.838378
\end{lstlisting}

\section{Impact of input parameters}

The results presented in the previous section have been obtained with the following input parameters (file HPC4E.txt):

\begin{cverbatim}
3 Dimension
"MAILLAGES/florian3D" Nom du maillage
.H (.dat or .model)
0 p-adaptivity (no=0)
1 Ordre des elements
0 transient
0.01 Simulation Time
0.01 Time Step for the snapshots
0 HDG formulation (1=yes, 0=no)
10 Plane wave ...
1
1
40 30 50
430 30 50
680 1380 50
1
2
0 Do you want curved elements
1 Do you want to use mumps with symmetry
0 seismograms 
1 Type de CLA (1 pour du/dn=iomega u)
.false.
0 type RTM
0 Analytic Solution (0=no)
\end{cverbatim}

\subsection{Scalability}

To show the impact of input parameters, we can change the elements order in HPC4E.txt (\textit{Ordre des elements}) or refine the mesh (\texttt{tail MAILLAGES/florian3D.1.ele} will show the command). Beware of the total execution time, it takes into account the time spent to write in a file (time in function \texttt{sub\_time\_loop}).

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements, {\bf{Simulation time=0.001}} (Time Step=0.001) - {\bf{without writting in sub\_time\_loop}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&  min 	& max	 & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				&  .7407	& .760 	& .746									&  	143.738 & 143.786	& 143.758 \\ \hline
 		8		                 				& .464	& .575  	& .512									&  	182.481 & 182.598 	& 182.554	\\ \hline		
 		16		                 				&  .142	& .174	 & .152									&  	215.348 & 215.398 	&  215.365\\ \hline
		 32		                 				& 1.288	&  5.884  & 4.011		  						        &  	293.542 &  293.639	& 293.619	\\ \hline
		64		                 				& 1.466	& 2.401 & 1.972									&  	360.580 & 360.936	& 360.886\\ \hline
\end{tabular}
}
\label{tab:1}
\end{table}

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements,  {\bf{Simulation time=0.01}} (Time Step=0.01)  - {\bf{without writting in sub\_time\_loop}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				& 5.504 & 5.543 & 5.525										& 156.409 & 156.481 & 156.438 	\\ \hline
 		8		                 				& 2.652  &  2.686 &  2.665									&   175.770  &  175.813 &  175.786	\\ \hline
 		16		                 				&  1.612 & 1.660  & 1.631 									& 217.008 &  217.055 &  217.026	\\ \hline
		 32		                 				& .924 & .948   & .940										& 292.669  &  292.749 & 292.730	\\ \hline
		64		                 				& .430 & .444  & .438										&  356.742  & 357.0359 &  357.001	\\ \hline
\end{tabular}
}
\label{tab:2}
\end{table}

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements,    {\bf{Simulation time=0.1}} (Time Step=0.1) - {\bf{without writting in sub\_time\_loop}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				& 55.325	& 55.344 & 55.333									&  194.500 & 194.522 & 194.512	\\ \hline
 		8		                 				& 26.804	& 26.843  & 26.820									&  200.059 & 200.105 & 200.083	\\ \hline
 		16		                 				&  17.671	&  17.710 & 17.688									&  233.131 & 233.162 &  233.145	\\ \hline
		 32		                 				& 9.779 & 9.803  & 9.794										&  301.450 & 301.533 & 301.514	\\ \hline
		64		                 				& 4.561 & 4.576  & 4.570										&  363.928 & 364.221 & 364.177	\\ \hline
\end{tabular}
}
\label{tab:3}
\end{table}

\subsection{Elements Order Variation}

Table \ref{tab:5} shows the total time and the time spent in function \texttt{sub\_loop\_time} for different elements orders, with 4 MPI processes. The times are obtained with the fortran function \texttt{CPU\_TIME}\footnote{Returns the CPU time, in seconds. If CPU\_TIME is used without any setting, the function returns the total user and system time for the current process.}. 
%The total execution time is obtained with \texttt{usr/bin/time}\footnote{Returns the elapsed real time between invocation and termination. The command is used with the -p option and the output format \%e (=elapsed real time in seconds).}.

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: 859247 elements,  Simulation time=0.001 (Time Step=0.001) - {\bf{without writting in sub\_time\_loop}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Elements order}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg							& 	min & max & avg			\\ \hline \hline
 		1		                 				& 1.273 & 1.374 & 1.298								& 156.247  & 156.435 &  156.309	\\ \hline
 		2		                 				& 3.384	&  3.390  & 3.388							&  171.825 & 171.855 & 171.843		\\ \hline
 		3		                 				& 59.873 	 &  61.902	& 61.060							&  578.569 & 579.954 & 579.564 	\\ \hline
		 4		                 				& 39.003	& 39.288  & 39.156							& 527.505  & 527.555 & 527.528		\\ \hline
		5		                 				& 	&  &											&  	 & 	& 		\\ \hline
\end{tabular}
}
\label{tab:5}
\end{table}

\subsection{Mesh refinement}

To refine the mesh, we launch the following command

\begin{cverbatim}
./tetgen -kna10000A florian3D.poly
\end{cverbatim}

\noindent and change the number in the option "-kna10000A".\\

Table \ref{tab:6} shows the total time and the time spent in function \texttt{sub\_loop\_time} for different mesh refinement, with 8 MPI processes.

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order=1, Simulation time=0.001 (Time Step=0.001)}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Mesh refinement}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg							& 	min & max & avg			\\ \hline \hline
 		10\ 000 (859\ 247 elt)		           	& .285 & .2973 &  .2917						&  175.0226 & 175.0522 & 175.041	\\ \hline
 		5\ 000 (1\ 704\ 052 elt)		                	& .8287 &  .8860 & .8590						&  348.888 & 379.0077 & 366.995	\\ \hline
 		2\ 500 (3\ 385\ 940 elt)	                		&  1.9168 &  1.9248	& 1.9214					&  708.008 & 707.994 &  707.993 \\ \hline
\end{tabular}
}
\label{tab:6}
\end{table}

%% Lancer avec 4 processus?? pour mieux comparer avec le reste

\section{Port hou10ni's kernel with BOAST}

This section presents how to install BOAST and use it to port Houdini's kernel.

\subsection{How to install BOAST}

\begin{cverbatim}
git clone https://github.com/Nanosim-LIG/boast.git
cd boast
gem build BOAST.gemspec
gem install --user-install BOAST-2.0.3.gem 
\end{cverbatim}

\subsection{Original version}

The kernel is the loop over the fluid inner elements in \texttt{lib/libtransient/sub\_time\_loop.f90}. The loop is written outside of the code, in a new subroutine as presented below.

\lstset{language=FORTRAN}
\begin{lstlisting}[caption=kernel.f90]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

  integer, intent(in) :: Nflu_inner
  integer, intent(in) :: Nflusol_inner
  integer, intent(in) :: nb_rhs
  integer, intent(in) :: afluSize
  integer, intent(in) :: pSize
  integer, intent(in) :: vecSize_1
  integer, intent(in) :: vecSize_2
  integer, intent(in) :: matSize
  INTEGER, INTENT(in), DIMENSION(vecSize_1,vecSize_2) :: idx_vec_flu
  INTEGER, INTENT(in), DIMENSION(matSize) :: idx_mat_flu
  REAL(kind=dp), INTENT(inout), DIMENSION(pSize,nb_rhs) :: P_new
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_inter
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_old
  REAL(kind=dp), INTENT(in), DIMENSION(afluSize) :: A_flu

  INTEGER :: I,J,I_tmp1,I_tmp2, I1, I2, I3, I4
  REAL(kind=dp), DIMENSION(20,nb_rhs) :: P_aux

  DO I=1,Nflu_inner+Nflusol_inner
    I_tmp1=1
    DO J=1,idx_vec_flu(1,I)
      I1=idx_vec_flu(2*J,I)
      I2=idx_vec_flu(2*J+1,I)
      I_tmp2=I_tmp1+I2-I1
      P_aux(I_tmp1:I_tmp2,1:nb_rhs)=P_inter(I1:I2,1:nb_rhs)
      I_tmp1=I_tmp2+1
    END DO
    I1=idx_vec_flu(2,I)
    I2=idx_vec_flu(3,I)
    I3=idx_mat_flu(I)
    I4=idx_mat_flu(I+1)-1

   P_new(I1:I2,1:nb_rhs)=MATMUL(RESHAPE(A_flu(I3:I4), (/I2-I1+1, I_tmp2/)),P_aux(1:I_tmp2,1:nb_rhs))-P_old(I1:I2,1:nb_rhs)
  END DO

END SUBROUTINE fluid_inner_elt_ref
\end{lstlisting}

\begin{lstlisting}[caption=sub\_time\_loop.f90]
SUBROUTINE sub_time_loop
...
CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
...
END SUBROUTINE 
\end{lstlisting}

To have an idea of the time spent in the kernel, table \ref{tab:7} shows the time of the entire function \texttt{sub\_time\_loop} (without the writing at the end) compared to the time spent in the kernel. Simulation time=0.01.

\begin{table}[H]
\centering
\caption{Time spent in the kernel (loop over fluid elements in {\tt sub\_time\_loop}), only one run - with an allocation at each step - {\bf{Avec l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|} \hline 
 {{\bf{Process ID}}} 	& {{\bf{Time in \texttt{sub\_time\_loop} }}}	& {{\bf{Time in \texttt{fluid\_inner\_elt} }}}\\ \hline \hline
		0		&   5.388336							&  5.1683220	(95,9\%)		\\ \hline 
 		1		&  5.352341							&  5.108320	(95,4\%)		\\ \hline
 		2		 &  5.360336							&  5.1203189	(95,5\%)		\\ \hline
		3		 &  5.412338 							&  5.1843260	(95,8\%)		\\ \hline	
\end{tabular}
}
\label{tab:7}
\end{table}

\begin{table}[H]
\centering
\caption{Time spent in the kernel (loop over fluid elements in {\tt sub\_time\_loop}), only one run - without allocation at each step - {\bf{Avec l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|} \hline 
 {{\bf{Process ID}}} 	& {{\bf{Time in \texttt{sub\_time\_loop} }}}	& {{\bf{Time in \texttt{fluid\_inner\_elt} }}}\\ \hline \hline
		0		&   4.776291							&  4.54828700	(95,2\%)		\\ \hline 
 		1		&  4.74029							&  4.5162959	(95,3\%)		\\ \hline
 		2		 &  4.744293							&  4.5242530	(95,4\%)		\\ \hline
		3		 & 4.768295 							&  4.53228600	(95\%)		\\ \hline	
\end{tabular}
}
\label{tab:8}
\end{table}

\subsection{Reference version}

We generate the same kernel as a reference version (KRef ) and run it with BOAST in order to compare the results in term of correctness and performances. This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST. A reference version should follow the template bellow.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef\_template.rb]
def KRef
  lang = BOAST::get_lang
  BOAST::set_lang(BOAST::FORTRAN)
  kernel = BOAST::CKernel::new
  kernel_name = "function_ref"

  # ARGUMENTS OF KERNEL HERE
  # FOR EXAMPLE : 
  geocode = BOAST::Int("geocode", :dir => :in )
  u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )

   p = BOAST::Procedure::new(kernel_name, [geocode,u])

   kernel.code.print <<EOF
  # ORIGINAL C or FORTRAN CODE HERE
EOF

  kernel.procedure = p
  BOAST::set_lang(lang)
  return kernel
end

geocode = ...
n01 = ...
u = ...
k = KRef
stats = k.run(geocode, n01, u)
puts "#{k.procedure.name}: #{stats[:duration]*1.0e3}"
\end{lstlisting}

We wrote the reference version a bite differently as shown bellow.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef.rb]
class KRef
 attr_reader :kernel

 def initialize(options)
  # Parameters 
  @Nflu_inner    = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner = Int("Nflusol_inner", :dir => :in )
  @nb_rhs        = Int("nb_rhs", :dir => :in )
  @afluSize      = Int("afluSize", :dir => :in )
  @pSize         = Int("pSize", :dir => :in )
  @vecSize_1     = Int("vecSize_1", :dir => :in )
  @vecSize_2     = Int("vecSize_2", :dir => :in )
  @matSize       = Int("matSize", :dir => :in )
  @idx_vec_flu   = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu   = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new         = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter       = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old         = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu         = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
 end

 def generate
  codeF = File::read("./kernel.f90")

  push_env(:lang => FORTRAN) {
     @kernel = CKernel:: new()
     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
     get_output.print codeF
  }
   return @kernel
 end

end
\end{lstlisting}

The file run\_ref.rb is used to launch the kernel.

\begin{lstlisting}[caption=run\_ref.rb]
require 'BOAST'
require './KRef.rb'
require 'narray_ffi'

include BOAST

 set_lang(FORTRAN)

# Creating ref kernel
k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
k = KRef::new(k_ref_params)
k.generate
puts k.kernel
k.kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )

# Input parameters 
# found in the directory fluid_inner_elt_ref/sub_time_loop_1/ for example (created by the user)
inputs = k.kernel.load_ref_inputs()
puts "** inputs =", inputs

# Launch the kernel fluid_inner_elt_ref
# As many key as directories in 
inputs.each_key { |key|
  puts k.kernel.run(*(inputs[key]))
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

puts "Testing done\n"
\end{lstlisting}

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, &
&vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

  integer, intent(in) :: Nflu_inner
  integer, intent(in) :: Nflusol_inner
  integer, intent(in) :: nb_rhs
  integer, intent(in) :: afluSize
  integer, intent(in) :: pSize
  integer, intent(in) :: vecSize_1
  integer, intent(in) :: vecSize_2
  integer, intent(in) :: matSize
  INTEGER, INTENT(in), DIMENSION(vecSize_1,vecSize_2) :: idx_vec_flu
  INTEGER, INTENT(in), DIMENSION(matSize) :: idx_mat_flu
  REAL(kind=dp), INTENT(inout), DIMENSION(pSize,nb_rhs) :: P_new
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_inter
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_old
  REAL(kind=dp), INTENT(in), DIMENSION(afluSize) :: A_flu

  !! Variables declaration
  INTEGER :: I,J,I_tmp1,I_tmp2, I1, I2, I3, I4
  REAL(kind=dp), DIMENSION(20,nb_rhs) :: P_aux

  DO I=1,Nflu_inner+Nflusol_inner
    I_tmp1=1
    DO J=1,idx_vec_flu(1,I)
      I1=idx_vec_flu(2*J,I)
      I2=idx_vec_flu(2*J+1,I)
      I_tmp2=I_tmp1+I2-I1
      P_aux(I_tmp1:I_tmp2,1:nb_rhs)=P_inter(I1:I2,1:nb_rhs)
      I_tmp1=I_tmp2+1
    END DO
    I1=idx_vec_flu(2,I)
    I2=idx_vec_flu(3,I)
    I3=idx_mat_flu(I)
    I4=idx_mat_flu(I+1)-1

   P_new(I1:I2,1:nb_rhs)=MATMUL(RESHAPE(A_flu(I3:I4), (/I2-I1+1, I_tmp&
&2/)),P_aux(1:I_tmp2,1:nb_rhs))-P_old(I1:I2,1:nb_rhs)
  END DO

END SUBROUTINE fluid_inner_elt_ref

** inputs =
{"./fluid_inner_elt_ref/sub_time_loop_1"=>[210025, 0, 1, NArray.int(2577732): 
[ 5, 1, 4, 538597, 538600, 319233, 319236, 363785, 363788, 665493, ... ], NArray.int(214812): 
[ 1, 81, 161, 209, 289, 369, 449, 529, 609, 689, 769, 849, 913, 993, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(8476280): 
[ 1.1026e-18, 9.61982e-17, 1.0165, 9.2055e-17, 9.13418e-19, 9.01442e-17, ... ], 16952560, 879732, 12, 214811, 214812]}

{:duration=>0.153594073}
\end{lstlisting}

\subsection{BOAST version}

\lstset{language=RUBY}
\begin{lstlisting}[caption=KBoast.rb]
class KBoast
 attr_reader :kernel

 def initialize(options)

  # Parameters 
  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
  @nb_rhs             = Int("nb_rhs", :dir => :in )
  @afluSize           = Int("afluSize", :dir => :in )
  @pSize              = Int("pSize", :dir => :in )
  @vecSize_1          = Int("vecSize_1", :dir => :in )
  @vecSize_2          = Int("vecSize_2", :dir => :in )
  @matSize            = Int("matSize", :dir => :in )
  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])

 end

 def generate

   push_env(:lang => FORTRAN) {

    @kernel = CKernel:: new()
    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)

    #..  FUNCTIONS ..#
    register_funccall("matmul") if get_lang == FORTRAN
    call_matmul = lambda{|x,y|
        return matmul(x,y)
    }

    #.. KERNEL ..#

    opn @kernel.procedure

    # local variables
    i = Variable::new('I', Int)
    j = Variable::new('J', Int)
    i_tmp1 = Variable::new('I_tmp1', Int)
    i_tmp2 = Variable::new('I_tmp2', Int)
    i1 = Variable::new('I1', Int)
    i2 = Variable::new('I2', Int)
    i3 = Variable::new('I3', Int)
    i4 = Variable::new('I4', Int)
    p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])

    decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux

    pr For(i,1,@Nflu_inner+@Nflusol_inner){
      pr i_tmp1 === 1
      pr For(j,1,@idx_vec_flu[1,i]){
        pr i1 === @idx_vec_flu[2*j,i]
        pr i2 === @idx_vec_flu[2*j+1,i]
        pr i_tmp2 === i_tmp1 + i2 - i1
        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs) 
        pr i_tmp1 === i_tmp2 +1
      }
      pr i1 === @idx_vec_flu[2,i]
      pr i2 === @idx_vec_flu[3,i]
      pr i3 === @idx_mat_flu[i]
      pr i4 === @idx_mat_flu[i+1] - 1
      reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
      pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
    }
    close @kernel.procedure
   }

   return @kernel
 end
end
\end{lstlisting}

\noindent The file run\_boast.rb is used to launch the BOAST version of the kernel. We run 3 times the kernel (repeat). \\

\begin{lstlisting}[caption=run\_boast.rb]
require 'BOAST'
require './KBoast.rb'
require 'narray_ffi'
include BOAST

k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}

set_default_real_size(4)
set_lang(FORTRAN)
stats = []
repeat = 5

# Creating boast kernel
k = KBoast::new(k_boast_params)
k.generate
puts k.kernel
k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )

# Get input and output parameters # found in the directory fluid_inner_elt_boast/
inputs = k.kernel.load_ref_inputs()
outputs = k.kernel.load_ref_outputs()

# Launch the kernel
inputs.each_key { |key|
  repeat.times {
    stats.push k.kernel.run(*(inputs[key]))
  }
  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
}

stats.sort_by! { |a| a[:duration] }
p stats
stats = stats.first

# Output
puts "#{k.kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
\end{lstlisting}

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, afluSize, pSize, vecSize_1, &
&vecSize_2, matSize, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
  integer, parameter :: wp=kind(1.0d0)
  integer(kind=4), intent(in) :: Nflu_inner
  integer(kind=4), intent(in) :: Nflusol_inner
  integer(kind=4), intent(in) :: nb_rhs
  ... 
END SUBROUTINE fluid_inner_elt_boast

{"P_new"=>0.0}  # No difference with the output

[{:duration=>0.011553553000000001, :start=>-8096482666880853504, :end=>-8084929113880853504}, {:duration=>0.011923121, :start=>-8122491033880853504, :end=>-8110567912880853504}, {:duration=>0.013262399000000001, :start=>-8135791916880853504, :end=>-8122529517880853504}, {:duration=>0.013991319, :start=>-8110525936880853504, :end=>-8096534617880853504}, {:duration=>0.014984313, :start=>-8150824959880853504, :end=>-8135840646880853504}]
fluid_inner_elt_boast: optim_nested = 3  0.011553553000000001 s ->  11.553553 ms
\end{lstlisting}

\subsection{Optimisations}

The following code shows how we optimized the kernel with loop unrolling and using OpenMP.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KBoast.rb]
class KBoast
 attr_reader :kernel

 def initialize(options)

  @opts = {:optim_nested => 1, :optim_main => 1}
  @opts.update(options)

  # Parameters 
  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
  @nb_rhs             = Int("nb_rhs", :dir => :in )
  @afluSize           = Int("afluSize", :dir => :in )
  @pSize              = Int("pSize", :dir => :in )
  @vecSize_1          = Int("vecSize_1", :dir => :in )
  @vecSize_2          = Int("vecSize_2", :dir => :in )
  @matSize            = Int("matSize", :dir => :in )
  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])

 end

  def generate

   push_env(:lang => FORTRAN) {

   @kernel = CKernel:: new()
   @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)

    #..  FUNCTIONS ..#

    register_funccall("matmul") if get_lang == FORTRAN
    call_matmul = lambda{|x,y|
        return matmul(x,y)
    }

    register_funccall(:modulo) if get_lang == FORTRAN

    #.. KERNEL ..#

    opn @kernel.procedure

    # local variables
    i = Variable::new('I', Int)
    j = Variable::new('J', Int)
    i_tmp1 = Variable::new('I_tmp1', Int)
    i_tmp2 = Variable::new('I_tmp2', Int)
    i1 = Variable::new('I1', Int)
    i2 = Variable::new('I2', Int)
    i3 = Variable::new('I3', Int)
    i4 = Variable::new('I4', Int)
    p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])

    jmax = Variable::new('jmax', Int)
    imax = Variable::new('imax', Int)

    decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
    decl jmax,imax

    ##.. Nested loop optimization
    optim_1 = lambda { |unroll,i|
      loop_1 = lambda { |indice|
        pr i1 === @idx_vec_flu[2*indice,i]
        pr i2 === @idx_vec_flu[2*indice+1,i]
        pr i_tmp2 === i_tmp1 + i2 - i1
        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs)
        pr i_tmp1 === i_tmp2 + 1.to_var
      }

      pr jmax === @idx_vec_flu[1,i]-(unroll-1)
      f_nested = For(j,1,jmax, step: unroll){
        unroll.times { |k|
          loop_1[j+k]
        }
      }
      f_arr_nested = [f_nested]
      if unroll > 1 then
          f2_nested = For(j,(@idx_vec_flu[1,i]-modulo(@idx_vec_flu[1,i],unroll))+1, @idx_vec_flu[1,i]){
            loop_1[j]
          }
          f_arr_nested.push f2_nested
      end
      f_arr_nested
   }

    ##.. Main loop optimization
    optim_2 = lambda { |unroll|
      loop_2 = lambda { |indice|
        pr i_tmp1 === 1
        # Nested loop optimized - No optimization -> = optim_1[1,indice].each{ |f|  pr f }
        optim_1[@opts[:optim_nested],indice].each { |f_nested|
          pr f_nested
        }
        pr i1 === @idx_vec_flu[2,indice]
        pr i2 === @idx_vec_flu[3,indice]
        pr i3 === @idx_mat_flu[indice]
        pr i4 === @idx_mat_flu[indice+1] - 1
        reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
        pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
      }

      pr imax === @Nflu_inner+@Nflusol_inner-(unroll-1)
      f_main = For(i,1,imax, step: unroll){
        unroll.times { |k|
          loop_2[i+k]
        }
      }
      f_arr_main = [f_main]
      if unroll > 1 then
        f2_main = For(i,(@Nflu_inner+@Nflusol_inner-modulo(@Nflu_inner+@Nflusol_inner,unroll))+1, @Nflu_inner+@Nflusol_inner){
          loop_2[i]
        }
        f_arr_main.push f2_main
      end
      f_arr_main
    }

    ## Procedure
    optim_2[@opts[:optim_main]].each { |f_main|
        pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads]){
        pr OpenMP::For(:schedule => "static")
        pr f_main
        }
    }

    close @kernel.procedure
   }

   return @kernel
 end    
end 
\end{lstlisting}

\subsubsection{Results}

The results were obtained using a Macbook Pro Intel Core i7 and the GNU fortran compiler. Values correspond to the minimum value of 20. Optimization flag: -O2
When \texttt{unroll main} and \texttt{unroll nested} are equal to 1, this means no unrolling is done.

\begin{table}[ht!]
\centering
\caption{Time depending on the number of OpenMP threads used for the main loop. No loop unrollment.}
{\scriptsize
\begin{tabular}{|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	  &  {{\bf{\# Threads}}}    &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
					 		  & 	 	[1..6]   &  				\\ \hline \hline
ref  &         -  &  {\bf{83,769018}} \\ \hline
boast &     1 & 90,608141 \\ \hline
boast &      2 & 87,766917 \\  \hline
boast &    3 & 90,69217800 \\ \hline
boast &     4 & 87,29976 \\  \hline
boast &    5 & 84,527453 \\  \hline
boast &   6 & {\bf{81,92376500}} \\ \hline
\end{tabular}
}
\label{tab:1}
\end{table}

\begin{table}[ht!]
\centering
\caption{Statistics on time depending on the number of iterations unrolled for the main loop.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 &   {{\bf{unroll main}}}   & {{\bf{unroll nested}}}   &    \multirow{2}{*}{ {{\bf{\# Threads}}} }  &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
							 &    	[1..3]		 	    &   [1..10]	 		 & 							      &  				\\ \hline \hline
boast &    1 &  &  & 90,608141 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 1 & 1 & 88,888357 \\  \cline{1-2}  \cline{5-5}
boast &    3 &  & & 89,98507500 \\ \hline \hline

boast &   1 &  &  & 88,516961 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 2 & 1 & 87,892416 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 86,117465 \\ \hline \hline

boast &    1 &  &  & 86,232686 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 3 & 1 & 86,40083400 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 82,334101 \\ \hline \hline

boast &    1 &  &  & 88,345071000 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 4 & 1 & 82,89916300 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 82,646548 \\ \hline \hline

boast &    1 & &  & 91,76817400 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 5 & 1 & 89,60873700 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 83,658755 \\ \hline \hline

boast &   1 & &  & 92,428135000 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 6 & 1 & 84,5453400 \\ \cline{1-2}  \cline{5-5}
boast &  3 &  &  & 87,82848900 \\ \hline \hline

boast &    1 &  &  & 85,81414200 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 7 & 1 & 89,83438 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  & & {\bf{81,77698900}}  \\ \hline \hline

boast &    1 &  &  & 85,221406 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 8 & 1 & 85,728041 \\ \cline{1-2}  \cline{5-5}
boast &   3 &  &  & 87,331661 \\ \hline \hline

boast &   1 &  &  & 88,48286500 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 9 & 1 & 87,05563300 \\ \cline{1-2}  \cline{5-5}
boast &    3 &  &  & 86,669093 \\ \hline \hline

boast &    1 &  &  & 92,284591 \\ \cline{1-2}  \cline{5-5}
boast &    2 & 10 & 1 & 85,477758 \\ \cline{1-2}  \cline{5-5}
boast &   3 & &  & 85,488478 \\ \hline
\end{tabular}
}
\label{tab:2}
\end{table}

\begin{table}[ht!]
\centering
\caption{Time depending on the number of iterations unrolled for the nested loop.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  {{\bf{unroll nested}}}   &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
						 		 &   		 					    &  	 		[1..10]				           & 	 						      &  				\\ \hline \hline
boast &     & 1 &  & 90,608141 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 2 &  & 88,516961 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 3 &  & 86,232686 \\ \cline{1-1}  \cline{3-3}  \cline{5-5} 
boast &   & 4 &  & 88,345071000 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &   1 & 5 & 1 & 91,76817400 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 6 &  & 92,42813500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 7 &  & 85,81414200 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 8 &  & {\bf{85,221406}} \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &    & 9 &  & 88,48286500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
boast &     & 10 &  & 92,284591 \\ \hline
\end{tabular}
}
\label{tab:3}
\end{table}

\begin{table}[ht!]
\centering
\caption{All values under the reference version.}
{\scriptsize
\begin{tabular}{|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  \multirow{2}{*}{ {{\bf{unroll nested}}} }  &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time}}} }  \\ 
						 		 &    							    &  	 							  & 	 						      &  				\\ \hline \hline
ref    &   &   &   &  0.083769018 \\ \hline \hline

boast &    3 & 7 & 1 & {\bf{81,77698900}} \\ \hline
boast &   2 & 3 & 5 & 82,006891 \\ \hline
boast &   3 & 4 & 3 & 82,17091900\\ \hline
boast &   3 & 3 & 1 & 82,334101 \\ \hline
boast &   3 & 4 & 1 & 82,646548 \\ \hline
boast &    2 & 4 & 5 & 82,809038 \\ \hline
boast &    2 & 4 & 4 & 82,81120700 \\ \hline
boast &    3 & 3 & 2 & 82,818522 \\ \hline
boast  &    1 & 5 & 4 & 82,866756 \\ \hline
boast &    2 & 4 & 1 & 82,89916300 \\ \hline
boast &    3 & 2 & 3 & 82,949804 \\ \hline
boast  &    1 & 3 & 4 & 82,994519 \\ \hline
boast &    1 & 2 & 3 & 83,047285 \\ \hline
boast &    3 & 7 & 2 & 83,08922 \\ \hline
boast  &    2 & 3 & 6 & 83,09479700 \\ \hline
boast &    3 & 4 & 6 & 83,137816 \\ \hline
boast &    3 & 1 & 3 & 83,15446500 \\ \hline
boast &    3 & 5 & 4 & 83,33983 \\ \hline
boast &    2 & 7 & 6 & 83,558517 \\ \hline
boast &    2 & 7 & 2 & 83,57951500 \\ \hline
\end{tabular}
}
\label{tab:4}
\end{table}

%\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
%TODO

\end{document}